---
title: "Supplementary Tutorial: A new effect size for meta-analysis of magnitude: lnM"
subtitle: "Worked examples + extra results from the simulation (figures + tables)"
author:
  - name: Shinichi Nakagawa
    attributes:
      corresponding: true
  - name: Ayumi Mizuno
  - name: Coralie Williams
  - name: Malgorzata Lagisz
  - name: Rose E. O'Dea
  - name: Yefeng Yang
  - name: Daniel W. A. Noble
  - name: Alistair M. Senior
  - name: Erick Lundgren
  - name: Santiago Ortega
format: 
  html:
    toc: true
    toc-location: left
    toc-depth: 3
    fig-dpi: 300
    toc-float: false
    toc-title: "**CONTENTS**"
    output-file: "index.html"
    embed-resources: true
    code-fold: false
    smooth-scroll: true
    anchor-sections: true
    code-tools: 
      source: false
    number-sections: false
    # bibliography: ./bib/ref.bib
    fontsize: "12"
    mainfont: "Roboto"
    monofont: "Roboto"
    # fontcolor: ""
    page-layout: full
    code-overflow: wrap
    df_print: paged
    theme: united
    code-line-numbers: false
    # grid:
    #   margin-width: 190px
execute:
  echo: true
  warning: false
  message: false
  cache: false
---

```{r}
#| label: setup
#| include: true
#| code-fold: true

# This chunk sets global chunk options, loads packages, and defines helper
# functions used throughout the supplement

knitr::opts_chunk$set(
  fig.align = "center",
  fig.width = 10,
  fig.height = 6,
  out.width = "100%"
)

suppressPackageStartupMessages({
  library(tidyverse)
  library(metafor)        # escalc(), rma.mv()
  library(here)           # robust file paths
  library(orchaRd)        # orchard_plot(), bubble_plot()
  library(kableExtra)     # HTML tables
  library(patchwork)      # combining ggplots   
  library(multcomp)        # multiple comparisons
})

# Reproducibility: SAFE uses Monte Carlo draws (bootstrap), so we fix a seed once.
set.seed(123)

# ---- lnM SAFE via orchaRd (internal functions) ----------------------------

get_lnM_safe_orchard <- function(m1, m2, s1, s2, n1, n2) {
  out <- orchaRd:::.safe_lnM_indep(m1, m2, s1, s2, n1, n2)

  tibble::tibble(
    yi_lnM_safe_orchard = out$lnM_SAFE,
    vi_lnM_safe_orchard = out$var_lnM_SAFE,
    draws_kept_orchard  = out$kept,
    draws_total_orchard = out$total,
    attempts_orchard    = out$attempts,
    status_orchard      = out$status
  )
}

# loading SAFE function
source(here("R", "SAFE_fun.R")) 

# ---- Folded normal helper (matches examples.R) ---------------------------
folded_norm <- function(mu, var) {
  sigma  <- sqrt(var)
  if (sigma == 0) return(c(point = abs(mu), var = 0))

  z      <- mu / sigma
  meanFN <- sigma * sqrt(2 / pi) * exp(-0.5 * z^2) +
    mu    * (2 * pnorm(z) - 1)
  varFN  <- mu^2 + var - meanFN^2

  c(point = meanFN, var = varFN)
}

# ---- lnM wrappers (match examples.R expectations) ------------------------
get_lnM_raw <- function(m1, m2, s1, s2, n1, n2) {
  out <- lnM_delta1_indep(m1, m2, s1, s2, n1, n2)   # returns c(point, var)
  tibble(
    yi_lnM_raw = out["point"],
    vi_lnM_raw = out["var"]
  )
}

get_lnM_safe <- function(m1, m2, s1, s2, n1, n2) {
  out <- safe_lnM_indep(m1, m2, s1, s2, n1, n2)     # returns list(point,var,kept,total)
  tibble(
    yi_lnM_safe = out$point,
    vi_lnM_safe = out$var,
    draws_kept  = out$kept,
    draws_total = out$total
  )
}




```

# Overview

This supplement has two parts:

Part I — Worked examples (Examples 1–2).

We demonstrate a complete, end-to-end workflow for quantifying effect magnitude using lnM. For each dataset, we first compute a conventional standardized mean difference (SMD) and compare results obtained using absolute SMD–based measures with those obtained using lnM.

We then estimate lnM using two approaches:

(i) a **Delta plug-in estimator** and
(ii) the **SAFE bootstrap bias-corrected point estimate (SAFE-BC)** which is designed to remain usable when the plug-in estimator is undefined (e.g., when $MS_B \le MS_W$).

Using these estimates, we fit multilevel meta-analytic models and meta-regressions and visualise results using the same plot types reported in the main manuscript.

<!-- End-to-end demonstration of the lnM workflow in real datasets: compute standardized mean differences, convert to folded-normal magnitudes, then compute lnM using (i) Delta plug-in and (ii) SAFE bootstrap bias-corrected point estimate (SAFE-BC). We then fit multilevel meta-analytic models and produce the same types of figures used in the manuscript. -->

Part II — Simulation reproduction (Figures + Tables).

We reproduce all simulation-based figures and tables reported in the manuscript using pre-computed simulation summaries and, where required, replicate-level outputs. These results characterise estimator performance in terms of bias, variance behaviour, empirical coverage, root mean square error (RMSE), and Monte Carlo standard error (MCSE).

We additionally provide compact tables summarising estimator feasibility, numerical stability, and SAFE bootstrap acceptance rates across simulation designs

<!--Load the simulation summary .rds and reproduce the manuscript/supplementary figures: point-estimate bias, variance relative bias, coverage, RMSE, and Monte-Carlo error diagnostics (MCSE). We also generate compact tables summarising acceptance rates, failure rates, and “worst offenders”.-->

# Part I — Worked examples

Part I illustrates how lnM is computed, analysed, and interpreted in real meta-analytic datasets, using two published examples with contrasting data structures.


## Example 1 — Almeida et al nutrition dataset

Almeida *et al.* examined whether experimental manipulations of nutrition produce consistently large deviations from control conditions across a range of physiological and performance traits. Because the biological interest lies in the magnitude of responses rather than their direction, this dataset provides a natural setting for comparing analyses based on absolute SMD measures with corresponding analyses based on lnM.

### Effect size calculations

```{r}
#| label: load-data example 1

# Load the dataset from CSV files

dat <- read_csv(here("data", "data_almeida.csv"), show_col_types = FALSE)
```

::: panel-tabset
## Conventional SMD (signed effect size)

```{r}
#|label: escalc-smd
dat <-  escalc(measure = "SMD", 
               m1i = m1i, 
               m2i = m2i, 
               sd1i = sd1i, 
               sd2i = sd2i, 
               n1i = n1i, 
               n2i = n2i,
               data = dat, append = TRUE,
               var.names = c("yi_d",   "vi_d"))
```

## Folded-normal magnitude of SMD 

```{r}
#| label: folded-normal-smd
dat <- dat %>%                               
  mutate(
    abs_d   = map2_dfr(yi_d,   vi_d, ~ {
      out <- folded_norm(.x, .y)
      tibble(yi_d_abs   = out["point"], vi_d_abs   = out["var"])
    })
  ) %>%        # flatten the tibbles
  unnest(c(abs_d))


```

## lnM via Delta plug-in estimator

```{r}
#| label: delta-lnM

dat <- dat %>% mutate(
  lnM_raw = pmap_dfr(
    list(m1i, m2i, sd1i, sd2i, n1i, n2i),
    get_lnM_raw)
) %>% unnest(lnM_raw)
```

## lnM via SAFE (single-fit parametric bootstrap)

::: pannel-tabset

```{r}
#| label: SAFE-lnM
# SAFE is stochastic, but the seed was set in the setup chunk for reproducibility
dat <- dat %>%
  dplyr::mutate(
    lnM_safe = purrr::pmap_dfr(
      list(m1i, m2i, sd1i, sd2i, n1i, n2i),
      get_lnM_safe
    )
  ) %>%
  tidyr::unnest(lnM_safe)

```

## lnM via SAFE (orchaRd)

```{r}
#| label: safe-lnm-orchard
lnM_safe_orchard_df <- purrr::pmap_dfr(
  list(dat$m1i, dat$m2i, dat$sd1i, dat$sd2i, dat$n1i, dat$n2i),
  get_lnM_safe_orchard
)

dat <- dplyr::bind_cols(dat, lnM_safe_orchard_df)


```


:::

:::

### How many lnM estimatescan be obtained?

The Delta plug-in estimator of lnM can be undefined for some effect sizes, particularly near boundary conditions. SAFE is designed to be more robust in these cases by relying on a single-fit parametric bootstrap. The table below summarises (i) how often each method yields a usable lnM point estimate and sampling variance, and (ii) how often SAFE provides valid estimates in cases where the Delta plug-in estimator fails.

::: callout-note
Percentages are computed relative to the total number of effect sizes in the dataset.
:::

Because lnM can be undefined when the between-group variance is not estimable, we first quantify how often each estimator returns usable lnM point estimates and sampling variances.

```{r}
#| label: lnM-availability-summary
#| message: false
#| warning: false
#| code-fold: true

stopifnot(exists("dat"), is.data.frame(dat))

n_total <- nrow(dat)

# ---- Helper counts -----------------------------------------------------------

summary_tbl <- tibble::tibble(
  Metric = c(
    "Total effect sizes",
    "Delta plug-in lnM: point estimate unavailable (undefined)",
    "Delta plug-in lnM: sampling variance unavailable",
    "SAFE lnM: point estimate unavailable",
    "SAFE lnM: sampling variance unavailable",
    "SAFE rescue: point estimate available when Delta plug-in is undefined",
    "SAFE rescue: sampling variance available when Delta plug-in is undefined"
  ),
  Count = c(
    n_total,
    sum(is.na(dat$yi_lnM_raw)),
    sum(is.na(dat$vi_lnM_raw)),
    sum(is.na(dat$yi_lnM_safe)),
    sum(is.na(dat$vi_lnM_safe)),
    sum(is.na(dat$yi_lnM_raw) & !is.na(dat$yi_lnM_safe)),
    sum(is.na(dat$vi_lnM_raw) & !is.na(dat$vi_lnM_safe))
  )
) %>%
  dplyr::mutate(
    Percent = dplyr::if_else(
      Metric == "Total effect sizes",
      NA_real_,
      100 * Count / n_total
    )
  )

# Headline numbers people care about
n_saved <- sum(!is.na(dat$yi_lnM_safe))
p_saved <- 100 * n_saved / n_total

cat(sprintf(
  "\n\n**SAFE produced defined lnM point estimates for %d/%d effect sizes (%.1f%%).**\n\n",
  n_saved, n_total, p_saved
))

summary_tbl %>%
  kableExtra::kable(
    digits = 1,
    caption = "Data availability for lnM estimates: Delta plug-in vs SAFE (counts and % of all effect sizes).",
    align = c("l", "r", "r")
  ) %>%
  kableExtra::kable_styling(full_width = FALSE)%>%
  kableExtra::scroll_box(width = "100%", height = "400px")
```

### Meta-analysis model

We fit the same multilevel random-effects structure for two magnitude metrics: (i) the folded-normal magnitude of SMD and (ii) lnM estimated with SAFE. 
In both cases, effect sizes are nested within studies (random intercepts for *Study* and for the within-study effect-size ID).

```{r}
#| label: subset-condition-example-1

# Restrict to the "Condition" subset used in the worked example

dat_cond <- dat %>% filter(Category == "Condition")
dim(dat_cond)
```

::: panel-tabset
## Absulte SMD (folded-normal magnitude)

This model summarises the typical magnitude of effects using the folded-normal transformation of the standardized mean difference (SMD). Rather than taking the absolute value of SMD directly, this approach uses the implied folded-normal distribution to obtain an approximately unbiased estimate of effect magnitude.

```{r}
# label: ma-abs-smd


ma_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML")

summary(ma_abs)
i2_ml(ma_abs)

# We added an "Overall" row to orchard plot for visualization
main <- mod_results(ma_abs, group = "Study")
main <- unclass(main)
main$mod_table$name <- "Overall" 
main$data$moderator <- "Overall"
class(main) <- "orchard"


p1 <- orchard_plot(main,  xlab = "abs(SMD)", 
                   group = "Study",
                   trunk.size = 0.6,
                   branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#CC6677") 

p1
```

## lnM (SAFE)

This model summarises the typical magnitude of effects using lnM estimated via SAFE. SAFE is used because lnM can be undefined under the Delta plug-in approach for some effect sizes, particularly when between-group variability is not larger than within-group variability.

```{r}
#| label: ma-lnM-safe

ma_safe <- rma.mv(
  yi   = yi_lnM_safe, 
  V    = vi_lnM_safe,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")

summary(ma_safe)
i2_ml(ma_safe)
ml_m1(ma_safe)


# Add an "Overall" row to the orchard object for visualization
main2 <- mod_results(ma_safe, group = "Study")
main2 <- unclass(main2)
main2$mod_table$name <- "Overall" 
main2$data$moderator <- "Overall"
class(main2) <- "orchard"


p2 <- orchard_plot(main2,  xlab = "lnM", 
                   group = "Study",
                   trunk.size = 0.6,
                   branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 

p2
```

```{r}
#| label: ma-lnM-safe-orchard
#| message: false
#| warning: false

# Meta-analysis using lnM estimated via SAFE (orchaRd implementation)
ma_safe_orchard <- rma.mv(
  yi     = yi_lnM_safe_orchard,
  V      = vi_lnM_safe_orchard,
  random = list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test   = "t"
)

summary(ma_safe_orchard)
i2_ml(ma_safe_orchard)
ml_m1(ma_safe_orchard)

# Add an "Overall" row to the orchard object for visualization
main2 <- mod_results(ma_safe_orchard, group = "Study")
main2 <- unclass(main2)
main2$mod_table$name <- "Overall"
main2$data$moderator <- "Overall"
class(main2) <- "orchard"

p2 <- orchard_plot(
  main2,
  xlab = "lnM (SAFE; orchaRd)",
  group = "Study",
  trunk.size = 0.6,
  branch.size = 5
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#117733")

p2

```


:::

::: callout-note
## Interpreting lnM on an SMD scale

For interpretability, we map $lnM$ onto an SMD-equivalent effect size, $d_{\mathrm{eq}}$, which is expressed in units of the within-group standard deviation. Using the large-sample approximation:

$d_{\mathrm{eq}}  \approx \sqrt{2} e^{\ln M}$

```{r}
#| label: lnm-to-deq-approx
#| message: false
#| warning: false

lnm_to_deq_approx <- function(model, label = "model") {
  b0_hat <- as.numeric(coef(model)[1])
  se_b0  <- sqrt(vcov(model)[1, 1])

  df_b0 <- tryCatch(as.numeric(model$ddf[1]), error = function(e) NA_real_)
  crit  <- if (is.finite(df_b0)) qt(0.975, df = df_b0) else qnorm(0.975)

  ci_lb <- b0_hat - crit * se_b0
  ci_ub <- b0_hat + crit * se_b0

  deq_hat <- sqrt(2) * exp(b0_hat)
  deq_ci  <- sqrt(2) * exp(c(ci_lb, ci_ub))

  cat(sprintf(
    "%s: d_eq (approx) = %.2f (95%% CI: %.2f to %.2f)\n",
    label, deq_hat, deq_ci[1], deq_ci[2]
  ))
}

lnm_to_deq_approx(ma_safe, label = "Overall model (ma_safe)")
```

Because the transformation is monotone, confidence intervals for  $d_{\mathrm{eq}}$ can be obtained by transforming the corresponding lnM interval endpoints
:::

### Meta-regression

We next examine whether variation in effect magnitude is associated with study-level covariates. Specifically, we regress the magnitude metric on *magnitude*, *duration*, and *Recovery* while retaining the same multilevel structure (random intercepts for *Study* and within-study effect-size ID). This allows a direct comparison of moderator patterns under a conventional absolute-value magnitude metric versus lnM.

:::: panel-tabset
## Absolute SMD (folded-normal magnitude)

```{r}
#| label: mod-abs-smd
mod_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  mods = ~ magnitude + duration + Recovery,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")
summary(mod_abs)

# bubble plot
p3 <- bubble_plot(mod_abs, mod = "Recovery", 
                  ylab = "abs(SMD)", 
                  xlab = "Recovery period", 
                  group = "Study") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#CC6677") 

p3
```

## lnM

```{r}
#| label: mod-lnM-safe
mod_safe <- rma.mv(
  yi   = yi_lnM_safe, 
  V    = vi_lnM_safe,
  mods = ~ magnitude + duration + Recovery,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")

summary(mod_safe)
# bubble plot
p4 <- bubble_plot(mod_safe, mod = "Recovery", 
                  ylab = "lnM", 
                  xlab = "Recovery period", 
                  group = "Study",
                  legend.pos = "bottom.left") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 
p4
```


::::

### Egger regression (small-study effects diagnostic)

We use an Egger-style meta-regression to assess whether estimated effect magnitudes systematically vary with study size (a pattern that can be consistent with small-study effects and publication bias). For lnM, using the usual standard-error-based predictor is problematic because lnM and its estimated standard error can be mechanically related. We therefore use a *sample-size-based* predictor derived from the harmonic mean sample size,

$n_0=\frac{n_1 n_2}{n_1+n_2}$,

and fit models using  $1/\sqrt{n_0}$ (an SE-analogue) and $1/n_0$ (a variance-analogue). These predictors depend only on sample sizes and avoid structural dependence on the observed lnM.

::: {.callout-important}
Because lnM estimates and their sampling standard errors are both functions of the same underlying variance components, they are structurally dependent. Using the sampling standard error as a predictor in Egger-type regressions would therefore violate the independence assumptions of the test.
:::

### Egger regression (publication bias test)

```{r}
#| label: ex1-egger-setup
# Create sample-size-based predictors using the harmonic mean sample size n0
dat_cond$n0 <- (dat_cond$n1i * dat_cond$n2i) / (dat_cond$n1i + dat_cond$n2i)
dat_cond$nSE <- 1 / sqrt(dat_cond$n0)
dat_cond$nV <- 1 / dat_cond$n0
```

:::: panel-tabset
## Absulte SMD

```{r}
# using abs(d)
egger_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  mods = ~ nSE,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")
summary(egger_abs)
# bubble plot
p5 <- bubble_plot(egger_abs, mod = "nSE", 
                  ylab = "abs(SMD)", 
                  xlab = "sqrt(inverse sample size)",
                  group = "Study") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#CC6677") 


p5
```

## lnM (1/sqrt(n0))

```{r}
#| label: ex1-egger-lnM-safe

egger_safe <- rma.mv(
  yi   = yi_lnM_safe, 
  V    = vi_lnM_safe,
  mods = ~ nSE,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")
summary(egger_safe)
# bubble plot
p6 <-bubble_plot(egger_safe, mod = "nSE", 
                 ylab = "lnM", 
                 xlab = "sqrt(inverse sample size)", 
                 group = "Study",
                 legend.pos = "bottom.right") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 

p6


```

## lnM (1/n0)

```{r}
#| label: ex1-egger-lnM-safe-2
# This specification uses the variance-analogue predictor 
egger_safe2 <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ nV,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")
summary(egger_safe2)
# bubble plot
bubble_plot(egger_safe2, mod = "nV", ylab = "lnM", 
            xlab = "Inverse sample size", group = "Study") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 
```

::: callout-note
## Mapping lnM to an SMD-equivalent effect size

For interpretability, we express lnM on a scale comparable to the standardized mean difference (SMD). Using the large-sample approximation ($d_{\mathrm{eq}}$), we get

```{r}
#| label: lnm-to-deq-approx-egger
lnm_to_deq_approx(egger_safe2, "Egger-style model (egger_safe2)")
```

indicating a substantially smaller average effect than conventional absolute-SMD estimates. 
:::
::::

### Sensitivity analysis

We repeated the lnM analyses after restricting effect sizes to comparisons with total sample size $n_1 + n_2 \geq 40$

```{r}
#| label: ex1-sensitivity-subset
dat_cond_sens <- dat_cond %>% filter((n1i + n2i) >= 40)
```

:::: panel-tabset
## lnM: meta-analysis

```{r}
#| label: ma-lnM-safe-sens
ma_safe_sens <- rma.mv(
  yi   = yi_lnM_safe, 
  V    = vi_lnM_safe,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond_sens,
  method = "REML",
  test = "t")

summary(ma_safe_sens)


i2_ml(ma_safe_sens)

orchard_plot(ma_safe_sens,  xlab = "lnM", group = "Study")
```
::: callout-note
## Mapping lnM to an SMD-equivalent effect size
```{r}
#| label: lnm-to-deq-approx-sens
lnm_to_deq_approx(ma_safe_sens, "Sensitivity model (ma_safe_sens)")
```
:::
## lnM meta-regression

::: panel-tabset
## Orginal meta-regression

```{r}
#| label: mod-lnM-safe-sens1
mod_safe2 <- rma.mv(
  yi   = yi_lnM_safe, 
  V    = vi_lnM_safe,
  mods = ~ magnitude + duration + Recovery + nSE,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")

summary(mod_safe2)
r2_ml(mod_safe2)

bubble_plot(mod_safe2, mod = "Recovery", 
            ylab = "lnM", 
            xlab = "Recovery period", 
            group = "Study",
            legend.pos = "bottom.left") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 
```

## Sensitivity meta-regression

```{r}
#| label: mod-lnM-safe-sens
mod_safe2 <- rma.mv(
  yi   = yi_lnM_safe, 
  V    = vi_lnM_safe,
  mods = ~ magnitude + duration + Recovery + nSE,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond_sens,
  method = "REML",
  test = "t")

summary(mod_safe2)
r2_ml(mod_safe2)

bubble_plot(mod_safe2, mod = "Recovery", 
            ylab = "lnM", 
            xlab = "Recovery period", 
            group = "Study",
            legend.pos = "bottom.left") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 
```
:::
::::

### Figure 3

```{r}
#| label: ex1-figure-3
# Combine panels into a 2 × 3 layout:
# Row 1: overall meta-analyses (absolute SMD vs lnM)
# Row 2: meta-regressions (absolute SMD vs lnM)
# Row 3: Egger-style diagnostics (absolute SMD vs lnM)
(p1 + p2) / (p3 + p4) / (p5 + p6) + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 16, face = "bold"))
```

## Example 2 — Kunc & Schmidt anthropogenic noise dataset

For this dataset, the underlying research question was whether exposure to anthropogenic noise alters animal behaviour and physiology relative to control (quiet) conditions. Each effect size contrasts a noise treatment with a control group, allowing the magnitude of noise-induced responses to be synthesised across taxa and response variables.

### Effect size calculations

```{r}
#| label: load-data example 2
#| 
# Workflow mirrors Example 1:
#   1) Load + basic cleaning
#   2) Compute SMD (here using unequal-variance SMDH)
#   3) Compute folded-normal magnitude of SMD
#   4) Compute lnM via Delta plug-in and via SAFE

# Load data
data<-read.csv(here("data", "kunc_2019.csv"), na.strings=c("","NA"))
data<-as.data.frame(data)


# 2) Basic cleaning: exclude rows with missing species information
data.2 <-data[which(data$species.latin!="NA"),]                


# 3) Ensure SD fields are numeric

data.2$sd.control<-as.numeric(as.character(data.2$sd.control))
data.2$sd.noise<-as.numeric(as.character(data.2$sd.noise))

# Exclude rows where sd.control is missing (cannot compute SMD/lnM)

data.2 <- data.2[which(!is.na(data.2$sd.control)),]




```

::: panel-tabset

##  Standardized mean difference with unequal variances (SMDH)
```{r}

# (Heteroscedastic SMD variant implemented in metafor)
dat <- escalc(measure="SMDH", 
              m1i=mean.control, sd1i=sd.control, n1i=sample.size.control, 
              m2i=mean.noise, sd2i=sd.noise,  n2i=sample.size.noise.1,
              data=data.2,
              append = TRUE,
              var.names = c("yi_d",   "vi_d"))

# Remove rows where effect size could not be computed
dat <- dat[which(!is.na(dat$yi_d)),]


```


## Folded-normal magnitude of SMD

```{r}
 
dat <- dat %>%                                   
  mutate(
    abs_d   = map2_dfr(yi_d,   vi_d,   ~ {
      out <- folded_norm(.x, .y)
      tibble(yi_d_abs   = out["point"], vi_d_abs   = out["var"])
    })) %>%
  unnest(abs_d)


```

## lnM via Delta plug-in estimator
```{r}
# 6) 
dat <- dat %>% mutate(
  lnM_raw = pmap_dfr(
    list(mean.control, mean.noise, sd.control, sd.noise, sample.size.control, sample.size.noise.1),
    get_lnM_raw)
) %>% unnest(lnM_raw)


```

## lnM via SAFE (single-fit parametric bootstrap)
```{r}

dat <- dat %>% mutate(
  lnM_safe = pmap_dfr(
    list(mean.control, mean.noise, sd.control, sd.noise, sample.size.control, sample.size.noise.1),
    get_lnM_safe)
) %>% unnest(lnM_safe)
```


:::


### How many lnM can be obtained?

```{r}
#| label: rescued-summary_ex2
#| message: false
#| warning: false
#| code-fold: true
#| 
stopifnot(exists("dat"), is.data.frame(dat))

n_total <- nrow(dat)

summary_tbl <- tibble::tibble(
  Metric = c(
    "Total effect sizes",
    "Delta plug-in lnM: point estimate unavailable (undefined)",
    "Delta plug-in lnM: sampling variance unavailable",
    "SAFE lnM: point estimate unavailable",
    "SAFE lnM: sampling variance unavailable",
    "SAFE rescue: point estimate available when Delta plug-in lnM is undefined",
    "SAFE rescue: sampling variance available when Delta plug-in lnM is undefined"
  ),
  Count = c(
    n_total,
    sum(is.na(dat$yi_lnM_raw)),
    sum(is.na(dat$vi_lnM_raw)),
    sum(is.na(dat$yi_lnM_safe)),
    sum(is.na(dat$vi_lnM_safe)),
    sum(is.na(dat$yi_lnM_raw) & !is.na(dat$yi_lnM_safe)),
    sum(is.na(dat$vi_lnM_raw) & !is.na(dat$vi_lnM_safe))
  )
) %>%
  dplyr::mutate(
    Percent = dplyr::if_else(
      Metric == "Total effect sizes",
      NA_real_,
      100 * Count / n_total
    )
  )

# Headline numbers (reported in text)
n_saved <- sum(!is.na(dat$yi_lnM_safe))
p_saved <- 100 * n_saved / n_total

cat(sprintf(
  "\n\n**SAFE produced usable lnM point estimates for %d/%d effect sizes (%.1f%%).**\n\n",
  n_saved, n_total, p_saved
))

summary_tbl %>%
  kableExtra::kable(
    digits = 1,
    caption = "Data availability for lnM estimates in the Kunc & Schmidt dataset: Delta plug-in versus SAFE (counts and % of all effect sizes).",
    align = c("l", "r", "r")
  ) %>%
  kableExtra::kable_styling(full_width = FALSE)
```

### Meta-analysis model

```{r}
#| label: ex2-filter-taxa
# Restrict to the focal taxa used in the worked example
# (exclude groups with very sparse representation in this dataset)
dat<-dat[which(dat$taxon.for.plot!="reptilia" & 
                 dat$taxon.for.plot!="mollusca" & 
                 dat$taxon.for.plot!="arthropoda"),]
```

::: panel-tabset
## Absolute SMD (folded-normal magnitude)

```{r}
#|label: ma-abs-smd-ex2
ma_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(ma_abs)
i2_ml(ma_abs)


# Add an "Overall" row to the orchard object for visualization

main <- mod_results(ma_abs, group = "study")
main <- unclass(main)
main$mod_table$name <- "Overall" 
main$data$moderator <- "Overall"
class(main) <- "orchard"

p7 <- orchard_plot(main,  xlab = "abs(SMD)", 
                   group = "study",
                   trunk.size = 0.6,
                   branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#CC6677") 

p7
```

## lnM

```{r}
#label: ma-lnM-safe-ex2
ma_safe <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(ma_safe)
i2_ml(ma_safe)

# Add an "Overall" row to the orchard object for visualization
main2 <- mod_results(ma_safe, group = "study")
main2 <- unclass(main2)
main2$mod_table$name <- "Overall" 
main2$data$moderator <- "Overall"
class(main2) <- "orchard"


p8 <- orchard_plot(main2,  xlab = "lnM", 
                   group = "study",
                   trunk.size = 0.8,
                   branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 


p8
```
::: callout-note
## Mapping lnM to an SMD-equivalent effect size
```{r}
#| label: lnm-to-deq-approx-ex2
lnm_to_deq_approx(ma_safe, "Overall model (ma_safe)")
```
:::
:::

### Meta-regression

::: panel-tabset
## Absolute SMD (folded-normal magnitude)

```{r}
#| label: mod-abs-smd-ex2
# Meta-regression using folded-normal |SMD| with taxonomic group as a categorical moderator.
# The "- 1" removes the intercept so each taxon level is estimated as its own mean.
mod_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  mods = ~ taxon.for.plot - 1,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(mod_abs)
# orchard plot
p9 <- orchard_plot(mod_abs, 
                   mod = "taxon.for.plot", 
                   xlab = "abs(SMD)", 
                   group = "study",
                   trunk.size = 0.6,
                   branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = rep("#CC6677", 4))

# multiple comparisons (Tukey-style contrasts among taxon means)
summary(glht(mod_abs, linfct=cbind(contrMat(rep(1,4), type="Tukey"))), test=adjusted("none"))

p9

```

## lnM

```{r}
#| label: mod-lnM-safe-ex2
mod_safe <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ taxon.for.plot -1,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(mod_safe)
# orchard plot
p10 <- orchard_plot(mod_safe, 
                    mod = "taxon.for.plot", 
                    xlab = "lnM", 
                    group = "study",
                    trunk.size = 0.6,
                    branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = rep("#117733", 4)) 
# multiple comparisons
summary(glht(mod_safe, linfct=cbind(contrMat(rep(1,4), type="Tukey"))), test=adjusted("none"))

p10
```
:::

### Egger regression (publication bias test)

```{r}
#| label: ex2-egger-setup
# Create sample-size-based predictors using the harmonic mean sample size n0

dat$n0 <- (dat$sample.size.control * dat$sample.size.noise.1
                 ) / (dat$sample.size.control + dat$sample.size.noise.1)
dat$nSE <- 1 / sqrt(dat$n0)
dat$nV <- 1 / dat$n0
```

::: panel-tabset
## Absolute SMD (folded-normal magnitude)

```{r}
#| label: ex2-egger-abs-smd
egger_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  mods = ~ nSE,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(egger_abs)
# bubble plot
p11 <- bubble_plot(egger_abs, mod = "nSE", 
                   ylab = "abs(SMD)", 
                   xlab = "sqrt(inverse sample size)", 
                   group = "study") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#CC6677") 

p11
```

## lnM (sqrt(1/n0))

```{r}
#| label: ex2-egger-lnM-safe
egger_safe <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ nSE,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(egger_safe)
# bubble plot
p12 <- bubble_plot(egger_safe, mod = "nSE", 
                   ylab = "lnM", 
                   xlab = "sqrt(inverse sample size)", 
                   group = "study",
                   legend.pos = "bottom.right") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 

p12
```

## lnM (1/n0)

```{r}
#| label: ex2-egger-lnM-safe-2
egger_safe2 <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ nV,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(egger_safe2)
# bubble plot
bubble_plot(egger_safe2, mod = "nV", ylab = "lnM", xlab = "Inverse sample size)", group = "study")
```
::: callout-note
## Mapping lnM to an SMD-equivalent effect size
```{r}
#| label: lnm-to-deq-approx-egger-ex2
lnm_to_deq_approx(egger_safe2, "Egger-style model (egger_safe2)")
```
:::
:::

### Sensitivity analysis

```{r}
#| label: ex2-sensitivity-subset

# Sensitivity subset: retain effect sizes with total sample size >= 40
dat_sens <- dat %>% filter((sample.size.control + sample.size.noise.1) >= 40)
```

:::: panel-tabset
## lnM: meta-analysis

```{r}
ma_safe_sens <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat_sens,
  method = "REML",
  test = "t")

summary(ma_safe_sens)


i2_ml(ma_safe_sens)

orchard_plot(ma_safe_sens,  xlab = "lnM", group = "study")
```
::: callout-note
## Mapping lnM to an SMD-equivalent effect size
```{r}
#| label: lnm-to-deq-approx-sens-ex2
lnm_to_deq_approx(ma_safe_sens, "Sensitivity model (ma_safe_sens)")
```
:::

## lnM meta-regression

::: panel-tabset
## Original lnM Egger's regression

```{r}
#| label: mod-lnM-safe-original-ex2
mod_safe2 <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ taxon.for.plot - 1 + nSE,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(mod_safe2)
# comparing
summary(mod_safe)

r2_ml(mod_safe2)

results <- mod_results(mod_safe2, 
                       group = "study", 
                       mod = "taxon.for.plot", 
                       at = list(nSE = c(0)))

# orchard plot
# probably closer to the truth
orchard_plot(results, 
                    mod = "taxon.for.plot", 
                    xlab = "lnM", 
                    group = "study",
                    trunk.size = 0.6,
                    branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = rep("#117733", 4)) 
```

## Sensitivity lnM Egger regression

```{r}
#| label: mod-lnM-safe-sens-ex2

mod_safe3 <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ taxon.for.plot - 1 + nSE,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat_sens,
  method = "REML",
  test = "t")
summary(mod_safe3)


r2_ml(mod_safe3)

results2 <- mod_results(mod_safe3, 
                       group = "study", 
                       mod = "taxon.for.plot", 
                       at = list(nSE = c(0)))

# orchard plot

orchard_plot(results2, 
             mod = "taxon.for.plot", 
             xlab = "lnM", 
             group = "study",
             trunk.size = 0.6,
             branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = rep("#117733", 4)) 

```
:::
::::

### Figure 4

```{r}
#| label: ex2-figure-4
# Combine panels into a 2 × 3 layout:
# Row 1: overall meta-analyses (absolute SMD vs lnM)
# Row 2: taxon meta-regressions (absolute SMD vs lnM)
# Row 3: Egger-style diagnostics (absolute SMD vs lnM)
(p7 + p8) / (p9 + p10) / (p11 + p12) + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 16, face = "bold"))
```

# Part II — Simulation reproduction (figures + tables)

Part II reproduces all simulation-based figures and tables reported in the manuscript, using pre-computed summary outputs and, where needed, replicate-level simulation files.

## Figures

### Load summary results

```{r}
#| label: sim-load
#| code-fold: true
# Load the pre-computed simulation summary used to reproduce figures and tables results

results <- readRDS( here::here("results", "lnM_summary_SAFEfun_2025-12-21.rds"))

# Create a compact facet label for plotting 
results <- results %>%
    mutate(facet_label = paste0(design, " n1=", n1, " n2=", n2))

facet_info <- results %>% distinct(facet_label, design, n1, n2)

balanced_ind <- facet_info %>%
    filter(design == "indep", n1 == n2) %>% arrange(n1) %>% pull(facet_label)
unbalanced_ind <- facet_info %>%
    filter(design == "indep", n1 != n2) %>% arrange(n1) %>% pull(facet_label)
paired_balanced <- facet_info %>%
    filter(design == "paired") %>% arrange(n1) %>% pull(facet_label)

new_levels <- c(balanced_ind, unbalanced_ind, paired_balanced)
results <- results %>% mutate(facet_label = factor(facet_label, levels = new_levels))

strip_labeller <- function(x) gsub(" n1=", "   n1=", gsub(" n2=", "   n2=", x))

# Shared plotting theme for all simulation figures

base_theme <- theme_bw(11) +
    theme(
      legend.title = element_text(),
      legend.position = "right",
      strip.background = element_rect(fill = "grey85", colour = "grey40"),
      strip.text = element_text(face = "plain"),
      panel.grid.minor = element_blank()
)
```

### Figure 1 — Bias of point estimates

```{r}
#| label: fig1-bias
#| fig-cap: "Figure 1. Point-estimate bias (estimate − true lnM). Dashed line indicates 0 bias."
#| code-fold: true
# Construct long-format data for plotting bias curves
bias_df <- dplyr::bind_rows(
    results %>% dplyr::select(theta, facet_label, delta_bias) %>%
      dplyr::rename(bias = delta_bias) %>% dplyr::mutate(estimator = "PI"),
    results %>% dplyr::select(theta, facet_label, safe_bias) %>%
      dplyr::rename(bias = safe_bias) %>% dplyr::mutate(estimator = "SAFE")
  )

  ggplot(bias_df, aes(theta, bias, colour = estimator, group = estimator)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_line(linewidth = 0.6) +
    facet_wrap(~ facet_label, ncol = 4, nrow = 3,
               labeller = labeller(facet_label = strip_labeller)) +
    labs(x = expression(theta),
         y = "Bias (estimate − true lnM)",
         colour = "estimator") +
    base_theme + 
  theme(legend.position = "bottom")#+
   # annotate("text", x = 0.7, y = max(bias_df$bias, na.rm = TRUE),
   #          label = "Target: 0 bias\nBias typically shrinks\nas θ and n increase",
   #          hjust = 0, vjust = 1, size = 3, colour = "grey25")
```

### Figure 2 — Relative bias of variance estimates

```{r}
#| label: fig2-relbias-var
#| fig-cap: "Figure 2. Relative bias of variance estimates (%). Dashed line indicates 0% (unbiased)."
#| code-fold: true

# Construct long-format data for plotting relative bias of variance estimates

rb_df <- dplyr::bind_rows(
    results %>% dplyr::select(theta, facet_label, relbias_delta) %>%
      dplyr::rename(relbias = relbias_delta) %>% dplyr::mutate(estimator = "Delta"),
    results %>% dplyr::select(theta, facet_label, relbias_safe) %>%
      dplyr::rename(relbias = relbias_safe)  %>% dplyr::mutate(estimator = "SAFE")
  )

  ggplot(rb_df, aes(theta, relbias, colour = estimator, group = estimator)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_line(linewidth = 0.6) +
    facet_wrap(~ facet_label, ncol = 4, nrow = 3,
               labeller = labeller(facet_label = strip_labeller)) +
    labs(x = expression(theta),
         y = "Relative bias of variance (%)",
         colour = "estimator") +
    base_theme + 
  theme(legend.position = "bottom")#+
    #annotate("text", x = 0.7, y = max(rb_df$relbias, na.rm = TRUE),
    #         label = "+: Var too large (conservative)\n−: Var too small (anti-conservative)",
    #         hjust = 0, vjust = 1, size = 3, colour = "grey25")
```

### Figure S1 — Coverage of nominal 95% intervals

```{r}
#| label: figS1-coverage
#| fig-cap: "Figure S1. Empirical coverage of nominal 95% intervals. Dashed line indicates 0.95."
#| code-fold: true
# Construct long-format data for plotting empirical coverage
cov_df <- dplyr::bind_rows(
    results %>% transmute(theta, facet_label, estimator = "PI",   cover = cover_delta),
    results %>% transmute(theta, facet_label, estimator = "SAFE", cover = cover_safe)
  )

  ggplot(cov_df, aes(theta, cover, colour = estimator, group = estimator)) +
    geom_hline(yintercept = 0.95, linetype = "dashed", colour = "grey50") +
    geom_line(linewidth = 0.6) +
    facet_wrap(~ facet_label, ncol = 4, nrow = 3,
               labeller = labeller(facet_label = strip_labeller)) +
    labs(x = expression(theta),
         y = "Empirical coverage",
         colour = "estimator") +
    base_theme + 
  theme(legend.position = "bottom")#+
    #annotate("text", x = 0.7, y = 0.99,
    #         label = "Target: 0.95\nBelow = under-coverage\nAbove = conservative",
    #        hjust = 0, vjust = 1, size = 3, colour = "grey25")

```

### Figure S2 — RMSE of point estimates

```{r}
#| label: figS2-rmse
#| fig-cap: "Figure S2. RMSE of point estimates. Lower values indicate better overall accuracy."
#| code-fold: true
# Construct long-format data for plotting RMSE curves
rmse_df <- dplyr::bind_rows(
    results %>% transmute(theta, facet_label, estimator = "PI",   rmse = rmse_delta),
    results %>% transmute(theta, facet_label, estimator = "SAFE", rmse = rmse_safe)
  )

  ggplot(rmse_df, aes(theta, rmse, colour = estimator, group = estimator)) +
    geom_line(linewidth = 0.6) +
    facet_wrap(~ facet_label, ncol = 4, nrow = 3,
               labeller = labeller(facet_label = strip_labeller)) +
    labs(x = expression(theta),
         y = expression(RMSE~"("~hat(lnM)~"\u2212"~lnM~")"),
         colour = "estimator") +
    base_theme + 
  theme(legend.position = "bottom")#+
   #annotate("text", x = 0.7, y = max(rmse_df$rmse, na.rm = TRUE),
    #         label = "RMSE combines\nbias + variance",
    #         hjust = 0, vjust = 1, size = 3, colour = "grey25")
```

### Reading in data for Monte Carlo (MC) standard error (SE)

```{r}
#| label: sim-mcse-load
#| code-fold: true

# Read replicate-level simulation outputs (one file per grid cell)
raw_dir <- here("raw_runs")
raw_files <- list.files(raw_dir, pattern = "^row_[0-9]{3}\\.rds$", full.names = TRUE)
if (!length(raw_files)) stop("No raw files found in: ", raw_dir)


# Extract grid-cell index from filename (e.g., row_012.rds -> 12)

row_index_from_path <- function(p) as.integer(sub("^row_([0-9]{3})\\.rds$", "\\1", basename(p)))

# Compute diagnostics for a single grid cell (one replicate-level data frame)


one_raw_diag <- function(raw_df) {
  
   # Defensive: remove replicate ID if present (not used in computations)
  if ("rep" %in% names(raw_df)) raw_df <- raw_df %>% dplyr::select(-rep)
  
  # Match the “ok_d” logic used in the simulation summary:
  # keep replicates where the Delta-method point & variance are defined and variance > 0
  ok_d <- is.finite(raw_df$delta_pt) & is.finite(raw_df$delta_var) & (raw_df$delta_var > 0)
  
  # If nothing survives the ok_d filter, return NAs
  if (!any(ok_d)) {
    return(data.frame(
      n_ok_d = 0L,
      n_ok_safe = 0L,
      RSE_VarMC_SAFEpt = NA_real_,
      sd_PI = NA_real_,
      sd_SAFE = NA_real_
    ))
  }
  
  df_ok <- raw_df[ok_d, , drop = FALSE]
  
  # Within the ok_d subset, check how many finite SAFE-BC point estimates exist
  ok_safe_pt <- is.finite(df_ok$safe_pt)
  n_ok_safe <- sum(ok_safe_pt)
  
  # Relative standard error (RSE) of the Monte Carlo variance estimate:
  # RSE(Var_hat) ≈ sqrt(2/(n-1)) for sample variance under normality
  RSE_varmc <- if (n_ok_safe >= 3) sqrt(2 / (n_ok_safe - 1)) else NA_real_
  
    # SD of point estimates (used to compute MCSE of bias later)
  # Since true_lnM is constant within a grid cell, sd(est - true) = sd(est).
  sd_PI   <- sd(df_ok$delta_pt, na.rm = TRUE)
  sd_SAFE <- sd(df_ok$safe_pt,  na.rm = TRUE)
  
  data.frame(
    n_ok_d = nrow(df_ok),
    n_ok_safe = n_ok_safe,
    RSE_VarMC_SAFEpt = RSE_varmc,
    sd_PI = sd_PI,
    sd_SAFE = sd_SAFE
  )
}
# Compute diagnostics for each grid cell (row_###.rds)
diag_list <- pbapply::pblapply(raw_files, function(f) {
  rid <- row_index_from_path(f)
  raw_df <- readRDS(f)
  d <- one_raw_diag(raw_df)
  d$row_id <- rid
  d
})

diag_df <- bind_rows(diag_list)
# Join diagnostics back onto the simulation summary
results2 <- results %>%
  mutate(row_id = seq_len(nrow(results))) %>%
  left_join(diag_df, by = "row_id")

```

### Figure S3 — MCSE of bias from raw_runs

```{r}
#| label: figS3-mcse-bias
#| fig-cap: "Figure S3. MCSE of bias (per grid cell) reconstructed from replicate-level files in raw_runs."
#| code-fold: true
# MCSE(bias) approximation:
#   MCSE(bias) ≈ sd(estimator) / sqrt(n_eff),
# where n_eff is the number of usable replicates in each grid cell.

results2 <- results2 %>%
  mutate(
    mcse_bias_PI_disk   = ifelse(is.finite(sd_PI)   & n_ok_d >= 2,    sd_PI   / sqrt(n_ok_d),     NA_real_),
    mcse_bias_SAFE_disk = ifelse(is.finite(sd_SAFE) & n_ok_safe >= 2, sd_SAFE / sqrt(n_ok_safe),  NA_real_)
  )

mc_long <- results2 %>%
  transmute(
    theta, facet_label,
    mcse_bias_PI   = mcse_bias_PI_disk,
    mcse_bias_SAFE = mcse_bias_SAFE_disk
  ) %>%
  pivot_longer(
    cols = c(mcse_bias_PI, mcse_bias_SAFE),
    names_to = "metric", values_to = "value"
  )

p_mc <- ggplot(mc_long, aes(theta, value, colour = metric, group = metric)) +
  geom_line() +
  facet_wrap(~ facet_label, ncol = 4) +
  labs(
    x = expression(theta),
    y = "MCSE of bias",
    title = "Monte-Carlo error diagnostics: MCSE of bias (per grid cell; from raw_runs)"
  ) +
  base_theme + 
  theme(legend.position = "bottom")

print(p_mc)
```

### Figure S4 — MCSE of mean variance estimates (from summary)

```{r}
#| label: figS4-mcse-varmean
#| fig-cap: "Figure S4. MCSE of mean variance estimates (from the simulation summary)."
#| code-fold: true

# Construct long-format data for plotting MCSE of mean variance estimates
var_mcse_df <- results %>%
  transmute(theta, facet_label,
            PI_var_MCSE   = mcse_varbar_delta,
            SAFE_var_MCSE = mcse_varbar_safe) %>%
  pivot_longer(cols = c(PI_var_MCSE, SAFE_var_MCSE),
               names_to = "estimator", values_to = "mcse") %>%
  mutate(estimator = recode(estimator,
                            PI_var_MCSE   = "PI (delta-var)",
                            SAFE_var_MCSE = "SAFE (safe-var)"))

p_var_mcse <- ggplot(var_mcse_df, aes(theta, mcse, colour = estimator, group = estimator)) +
  geom_line() +
  facet_wrap(~ facet_label, ncol = 4) +
  labs(x = expression(theta),
       y = "MCSE of mean(variance estimate)",
       colour = NULL,
       title = "Monte-Carlo error of variance estimators (MCSE of the mean)") +
  base_theme + 
  theme(legend.position = "bottom")

print(p_var_mcse)
```

## Tables for the supplement



### Table S1 — Method feasibility and stability

Table S1 summarises feasibility and numerical stability diagnostics for lnM estimation across simulation designs and sample-size combinations. All quantities are summarised across values of the true lnM (θ) within each design × (n₁, n₂) facet.
Medians describe typical behaviour across θ, whereas maxima highlight worst-case performance near parameter boundaries.


**Column definitions.**  
- **θ_min, θ_max**: Minimum and maximum true lnM values evaluated in the simulation grid for a given design × sample-size combination.  
- **delta_fail_median**: Median proportion of simulation replicates (across θ) for which the Delta plug-in lnM estimator was undefined (e.g. due to boundary violations).  
- **safe_fail_median**: Median proportion of simulation replicates for which SAFE failed to return a usable lnM estimate.  
- **SAFE_ok_median**: Median proportion of replicates in which SAFE returned a valid (accepted) bootstrap estimate.  
- **delta_cap_median**: Median proportion of replicates in which the Delta-method sampling variance was capped for numerical stability.  
- **delta_fail_max, safe_fail_max, delta_cap_max**: Worst-case (maximum) values of the corresponding metrics across θ, highlighting boundary regions where estimator behaviour deteriorates.

::: {.callout-note appearance="simple"}
```{r}
#| label: tabS1-feasibility
#| code-fold: true

tab1_stability <- results %>%
  group_by(design, n1, n2, facet_label) %>%
  summarise(
    # Range of θ (theta) values evaluated in each facet
    theta_min = min(theta, na.rm = TRUE),
    theta_max = max(theta, na.rm = TRUE),
    
    # Median failure rates across θ
    delta_fail_median = median(delta_fail_prop, na.rm = TRUE),
    safe_fail_median  = median(safe_fail_prop,  na.rm = TRUE),
    
    # Median SAFE convergence / acceptance rate
    SAFE_ok_median    = median(SAFE_ok_rate, na.rm = TRUE),
    
    # Median rate of Delta variance capping
    delta_cap_median  = median(delta_cap_rate, na.rm = TRUE),
    
    # Worst-case behaviour across θ
    delta_fail_max = max(delta_fail_prop, na.rm = TRUE),
    safe_fail_max  = max(safe_fail_prop,  na.rm = TRUE),
    delta_cap_max  = max(delta_cap_rate,  na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  arrange(factor(facet_label, levels = levels(results$facet_label))) %>%
  mutate(
    design = as.character(design),
    facet_label = as.character(facet_label)
  )

kable(tab1_stability, digits = 3,
      caption = "Table S1. Feasibility / stability diagnostics summarised across θ within each design × (n1,n2) facet.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))%>%
  kableExtra::scroll_box(width = "100%", height = "400px")
```
:::
### Table S2 — SAFE bootstrap acceptance (kept/tried)

Table S2 summarises the bootstrap acceptance rate of SAFE across simulation designs and sample-size combinations. 
The acceptance rate is defined as the proportion of bootstrap resamples retained (“kept”) after stability and feasibility checks, relative to the total number attempted.
Quantiles describe the distribution of acceptance rates across values of the true lnM (θ), highlighting both typical and worst-case performance.

**Column definitions.**  
- **q0, q25, q50, q75, q100**: Minimum, 25th percentile, median, 75th percentile, and maximum SAFE bootstrap acceptance rates across values of θ within each facet.  
- **mean**: Mean SAFE bootstrap acceptance rate across θ.  

::: {.callout-note appearance="simple"}
```{r}
#| label: tabS2-safe-acceptance
#| code-fold: true

tab2_accept <- results %>%
  group_by(design, n1, n2, facet_label) %>%
  summarise(
    q0  = quantile(boot_accept_prop, 0.00, na.rm = TRUE),
    q25 = quantile(boot_accept_prop, 0.25, na.rm = TRUE),
    q50 = quantile(boot_accept_prop, 0.50, na.rm = TRUE),
    q75 = quantile(boot_accept_prop, 0.75, na.rm = TRUE),
    q100= quantile(boot_accept_prop, 1.00, na.rm = TRUE),
    mean = mean(boot_accept_prop, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(factor(facet_label, levels = levels(results$facet_label)))

kable(tab2_accept, digits = 4,
      caption = "Table S2. SAFE acceptance rate (kept/tried) summarised across θ for each facet. Quantiles highlight worst-case acceptance.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))%>%
  kableExtra::scroll_box(width = "100%", height = "400px")
```
Acceptance rate is defined as the proportion of bootstrap resamples retained after SAFE’s feasibility and stability checks. Lower values indicate reduced effective bootstrap sample sizes, typically occurring near parameter boundaries or under small sample sizes.
:::
### Table S3 — MCSE diagnostics
Table S3 summarises Monte Carlo precision diagnostics for the simulation study.
Reported values correspond to the maximum Monte Carlo standard error (MCSE) observed across values of the true lnM (θ) within each design × (n₁, n₂) facet, providing a conservative assessment of simulation uncertainty.


**Column definitions.**  
- **mcse_bias_PI_max**: Maximum Monte Carlo standard error of the Delta-method lnM bias estimate across θ.  
- **mcse_bias_SAFE_max**: Maximum Monte Carlo standard error of the SAFE lnM bias estimate across θ.  
- **mcse_var_PI_max**: Maximum Monte Carlo standard error of the mean Delta-method variance estimate across θ.  
- **mcse_var_SAFE_max**: Maximum Monte Carlo standard error of the mean SAFE variance estimate across θ.  



::: {.callout-note appearance="simple"}
```{r}
#| label: tabS3-mcse
#| code-fold: true

has_mcse_bias <- all(c("mcse_bias_delta","mcse_bias_safe") %in% names(results))

tab3_mcse <- results %>%
  group_by(design, n1, n2, facet_label) %>%
  summarise(
    # bias MCSE (optional)
    mcse_bias_PI_max   = if (has_mcse_bias) max(mcse_bias_delta, na.rm = TRUE) else NA_real_,
    mcse_bias_SAFE_max = if (has_mcse_bias) max(mcse_bias_safe,  na.rm = TRUE) else NA_real_,
    
    # variance MCSE (these are in your results)
    mcse_var_PI_max    = max(mcse_varbar_delta, na.rm = TRUE),
    mcse_var_SAFE_max  = max(mcse_varbar_safe,  na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  arrange(factor(facet_label, levels = levels(results$facet_label)))

kable(tab3_mcse, digits = 5,
      caption = "Table S3. Monte-Carlo precision (max MCSE over θ within each facet).") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped","hover","condensed"))%>%
  kableExtra::scroll_box(width = "100%", height = "400px")

```
MCSE values quantify simulation uncertainty due to a finite number of Monte Carlo replicates; smaller values indicate more precise simulation estimates.
:::

# Session info

```{r}
#| label: session-info
sessionInfo()
```
