---
title: "Supplementary Tutorial: lnM"
subtitle: "Worked examples + extra results from the simulation (figures + tables)"
author: "TBA"
format: 
  html:
    toc: true
    toc-location: left
    toc-depth: 6
    toc-float: false
    toc-title: "**CONTENTS**"
    output-file: "index.html"
    embed-resources: true
    # code-fold: true
    code-tools: 
      source: false
    number-sections: false
    # bibliography: ./bib/ref.bib
    fontsize: "12"
    mainfont: "Roboto"
    monofont: "Roboto"
    # fontcolor: ""
    page-layout: full
    code-overflow: wrap
    df_print: paged
    theme: united
    code-line-numbers: false
    # grid:
    #   margin-width: 190px
execute:
  echo: true
  warning: false
  message: false
  cache: false
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  fig.align = "center",
  fig.width = 10,
  fig.height = 6,
  out.width = "100%"
)

suppressPackageStartupMessages({
  library(tidyverse)
  library(metafor)        # escalc(), rma.mv()
  library(here)
  library(orchaRd)
  library(kableExtra)
  library(patchwork)
  library(multcomp)
})

# loading SAFE function
source(here("R", "SAFE_fun.R")) 

# ---- Folded normal helper (matches examples.R) ---------------------------
folded_norm <- function(mu, var) {
  sigma  <- sqrt(var)
  if (sigma == 0) return(c(point = abs(mu), var = 0))

  z      <- mu / sigma
  meanFN <- sigma * sqrt(2 / pi) * exp(-0.5 * z^2) +
    mu    * (2 * pnorm(z) - 1)
  varFN  <- mu^2 + var - meanFN^2

  c(point = meanFN, var = varFN)
}

# ---- lnM wrappers (match examples.R expectations) ------------------------
get_lnM_raw <- function(m1, m2, s1, s2, n1, n2) {
  out <- lnM_delta1_indep(m1, m2, s1, s2, n1, n2)   # returns c(point, var)
  tibble(
    yi_lnM_raw = out["point"],
    vi_lnM_raw = out["var"]
  )
}

get_lnM_safe <- function(m1, m2, s1, s2, n1, n2) {
  out <- safe_lnM_indep(m1, m2, s1, s2, n1, n2)     # returns list(point,var,kept,total)
  tibble(
    yi_lnM_safe = out$point,
    vi_lnM_safe = out$var,
    draws_kept  = out$kept,
    draws_total = out$total
  )
}
```

# Overview

This supplement has two parts:

Part I — Worked examples (Examples 1–2).
End-to-end demonstration of the lnM workflow in real datasets: compute standardized mean differences, convert to folded-normal magnitudes, then compute lnM using (i) Delta plug-in and (ii) SAFE bootstrap bias-corrected point estimate (SAFE-BC). We then fit multilevel meta-analytic models and produce the same types of figures used in the manuscript.

Part II — Simulation reproduction (Figures + Tables).
Load the simulation summary .rds and reproduce the manuscript/supplementary figures: point-estimate bias, variance relative bias, coverage, RMSE, and Monte-Carlo error diagnostics (MCSE). We also generate compact tables summarising acceptance rates, failure rates, and “worst offenders”.


# Part I — Worked examples

## Example 1 — Almeida nutrition dataset

### Data + effect size

```{r}
# data - read in from CSV files

dat <- read_csv(here("data", "data_almeida.csv"), show_col_types = FALSE)

# effect size calculations 

# getting normal d
dat <-  escalc(measure = "SMD", 
               m1i = m1i, 
               m2i = m2i, 
               sd1i = sd1i, 
               sd2i = sd2i, 
               n1i = n1i, 
               n2i = n2i,
               data = dat, append = TRUE,
               var.names = c("yi_d",   "vi_d"))


# absolute d
dat <- dat %>%                               
  mutate(
    abs_d   = map2_dfr(yi_d,   vi_d,   ~ {
      out <- folded_norm(.x, .y)
      tibble(yi_d_abs   = out["point"], vi_d_abs   = out["var"])
    })
  ) %>%                                           # flatten the tibbles
  unnest(c(abs_d))


# lnM
dat <- dat %>% mutate(
  lnM_raw = pmap_dfr(
    list(m1i, m2i, sd1i, sd2i, n1i, n2i),
    get_lnM_raw)
) %>% unnest(lnM_raw)

# SAFE is stochastic so set a seed
set.seed(123)

dat <- dat %>% mutate(
  lnM_safe = pmap_dfr(
    list(m1i, m2i, sd1i, sd2i, n1i, n2i),
    get_lnM_safe)
) %>% unnest(lnM_safe)
```

### How many lnM can be obtained?

```{r}
#| message: false
#| warning: false

stopifnot(exists("dat"), is.data.frame(dat))

n_total <- nrow(dat)

summary_tbl <- tibble::tibble(
  Metric = c(
    "Total effect sizes",
    "Delta lnM: point estimate missing",
    "Delta lnM: sampling variance missing",
    "SAFE lnM: point estimate missing",
    "SAFE lnM: sampling variance missing",
    "Rescued by SAFE (point available when Delta missing)",
    "Rescued by SAFE (variance available when Delta missing)"
  ),
  Count = c(
    n_total,
    sum(is.na(dat$yi_lnM_raw)),
    sum(is.na(dat$vi_lnM_raw)),
    sum(is.na(dat$yi_lnM_safe)),
    sum(is.na(dat$vi_lnM_safe)),
    sum(is.na(dat$yi_lnM_raw) & !is.na(dat$yi_lnM_safe)),
    sum(is.na(dat$vi_lnM_raw) & !is.na(dat$vi_lnM_safe))
  )
) %>%
  dplyr::mutate(
    Percent = dplyr::if_else(
      Metric == "Total effect sizes",
      NA_real_,
      100 * Count / n_total
    )
  )

# Headline numbers people care about
n_saved <- sum(!is.na(dat$yi_lnM_safe))
p_saved <- 100 * n_saved / n_total

cat(sprintf(
  "\n\n**SAFE produced usable lnM point estimates for %d/%d effect sizes (%.1f%%).**\n\n",
  n_saved, n_total, p_saved
))

summary_tbl %>%
  kableExtra::kable(
    digits = 1,
    caption = "Data availability for lnM estimates: Delta plug-in vs SAFE (counts and % of all effect sizes).",
    align = c("l", "r", "r")
  ) %>%
  kableExtra::kable_styling(full_width = FALSE)
```

### Meta-analysis model 


```{r}
# filtering for just condition data
# just using a part of data

dat_cond <- dat %>% filter(Category == "Condition")
dim(dat_cond)
```


::: {.panel-tabset}

### Absulte SMD

```{r}

# meta-analysis using abs(d)

ma_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML")

summary(ma_abs)
i2_ml(ma_abs)

# adding overll to orchard plot
main <- mod_results(ma_abs, group = "Study")
main <- unclass(main)
main$mod_table$name <- "Overall" 
main$data$moderator <- "Overall"
class(main) <- "orchard"

p1 <- orchard_plot(main,  xlab = "abs(SMD)", 
                   group = "Study",
                   trunk.size = 0.6,
                   branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#CC6677") 

p1
```

### lnM

```{r}
# meta-analysis using safe

ma_safe <- rma.mv(
  yi   = yi_lnM_safe, 
  V    = vi_lnM_safe,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")

summary(ma_safe)
i2_ml(ma_safe)
ml_m1(ma_safe)


# adding overll to orchard plot
main2 <- mod_results(ma_safe, group = "Study")
main2 <- unclass(main2)
main2$mod_table$name <- "Overall" 
main2$data$moderator <- "Overall"
class(main2) <- "orchard"


p2 <- orchard_plot(main2,  xlab = "lnM", 
                   group = "Study",
                   trunk.size = 0.6,
                   branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 

p2
```

:::

### Meta-regression

::: {.panel-tabset}

### Absulte SMD

```{r}
# meta-regression abs(d)
mod_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  mods = ~ magnitude + duration + Recovery,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")
summary(mod_abs)
# bubble plot
p3 <- bubble_plot(mod_abs, mod = "Recovery", 
                  ylab = "abs(SMD)", 
                  xlab = "Recovery period", 
                  group = "Study") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#CC6677") 

p3
```

### lnM

```{r}
# meta-regression lnM
mod_safe <- rma.mv(
  yi   = yi_lnM_safe, 
  V    = vi_lnM_safe,
  mods = ~ magnitude + duration + Recovery,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")

summary(mod_safe)
# bubble plot
p4 <- bubble_plot(mod_safe, mod = "Recovery", 
                  ylab = "lnM", 
                  xlab = "Recovery period", 
                  group = "Study",
                  legend.pos = "bottom.left") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 
```

:::

### Egger regression (publication bias test)

```{r}
# Creaing a variable using n0
dat_cond$n0 <- (dat_cond$n1i * dat_cond$n2i) / (dat_cond$n1i + dat_cond$n2i)
dat_cond$nSE <- 1 / sqrt(dat_cond$n0)
dat_cond$nV <- 1 / dat_cond$n0
```


::: {.panel-tabset}

### Absulte SMD

```{r}
# using abs(d)
egger_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  mods = ~ nSE,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")
summary(egger_abs)
# bubble plot
p5 <- bubble_plot(egger_abs, mod = "nSE", 
                  ylab = "abs(SMD)", 
                  xlab = "sqrt(inverse sample size)",
                  group = "Study") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#CC6677") 


p5
```

### lnM (sqrt(1/n0))

```{r}
# using safe
egger_safe <- rma.mv(
  yi   = yi_lnM_safe, 
  V    = vi_lnM_safe,
  mods = ~ nSE,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")
summary(egger_safe)
# bubble plot
p6 <-bubble_plot(egger_safe, mod = "nSE", 
                 ylab = "lnM", 
                 xlab = "sqrt(inverse sample size)", 
                 group = "Study",
                 legend.pos = "bottom.right") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 

p6


```

### lnM (1/n0)

```{r}
# equivalent to fitting V rather than SE
egger_safe2 <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ nV,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")
summary(egger_safe2)
# bubble plot
bubble_plot(egger_safe, mod = "nSE", ylab = "lnM", 
            xlab = "sqrt(inverse sample size)", group = "Study") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 
```


:::


### Sensitivity analysis

```{r}
# sensitivity analysis - we want to get effect sizes where n1 + n2 = or > 40
dat_cond_sens <- dat_cond %>% filter((n1i + n2i) >= 40)
```


::: {.panel-tabset}

### lnM: meta-analysis

```{r}
ma_safe_sens <- rma.mv(
  yi   = yi_lnM_safe, 
  V    = vi_lnM_safe,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond_sens,
  method = "REML",
  test = "t")

summary(ma_safe_sens)
summary(ma_safe)

i2_ml(ma_safe_sens)
#ml_m1(ma_safe_sens)

orchard_plot(ma_safe_sens,  xlab = "lnM", group = "Study")
```

### lnM meta-regression

```{r}
# doing meta-regression again
# meta-regression lnM
mod_safe2 <- rma.mv(
  yi   = yi_lnM_safe, 
  V    = vi_lnM_safe,
  mods = ~ magnitude + duration + Recovery + nSE,
  random =  list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test = "t")

summary(mod_safe2)
r2_ml(mod_safe2)

bubble_plot(mod_safe2, mod = "Recovery", 
            ylab = "lnM", 
            xlab = "Recovery period", 
            group = "Study",
            legend.pos = "bottom.left") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 
```

:::

### Figure 3

```{r}
# making 2 x 3 panel fig
(p1 + p2) / (p3 + p4) / (p5 + p6) + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 16, face = "bold"))
```

## Example 2 — Kunc & Schmidt anthropogenic noise dataset

### Data + effect size

```{r}
data<-read.csv(here("data", "kunc_2019.csv"), na.strings=c("","NA"))
data<-as.data.frame(data)
data.2 <-data[which(data$species.latin!="NA"),]                ### exclude cases where no species were provided
str(data.2)

#  make sd.control and sd.noise numeric

data.2$sd.control<-as.numeric(as.character(data.2$sd.control))
data.2$sd.noise<-as.numeric(as.character(data.2$sd.noise))

# exclude NA in dat$sd.control

data.2 <- data.2[which(!is.na(data.2$sd.control)),]

# hetro-verion of SMD
dat <- escalc(measure="SMDH", 
              m1i=mean.control, sd1i=sd.control, n1i=sample.size.control, 
              m2i=mean.noise, sd2i=sd.noise,  n2i=sample.size.noise.1,
              data=data.2,
              append = TRUE,
              var.names = c("yi_d",   "vi_d"))

# fitering out NA for yi_d and vi_d
dat <- dat[which(!is.na(dat$yi_d)),]

dat <- dat %>%                                   # <- your existing data frame
  mutate(
    abs_d   = map2_dfr(yi_d,   vi_d,   ~ {
      out <- folded_norm(.x, .y)
      tibble(yi_d_abs   = out["point"], vi_d_abs   = out["var"])
    })
  ) %>%                                           # flatten the tibbles
  unnest(c(abs_d))

# lnM
dat <- dat %>% mutate(
  lnM_raw = pmap_dfr(
    list(mean.control, mean.noise, sd.control, sd.noise, sample.size.control, sample.size.noise.1),
    get_lnM_raw)
) %>% unnest(lnM_raw)

# SAFE is stochastic so set a seed
set.seed(123)

dat <- dat %>% mutate(
  lnM_safe = pmap_dfr(
    list(mean.control, mean.noise, sd.control, sd.noise, sample.size.control, sample.size.noise.1),
    get_lnM_safe)
) %>% unnest(lnM_safe)
```

### How many lnM can be obtained?

```{r}
#| label: rescued-summary
#| message: false
#| warning: false

stopifnot(exists("dat"), is.data.frame(dat))

n_total <- nrow(dat)

summary_tbl <- tibble::tibble(
  Metric = c(
    "Total effect sizes",
    "Delta lnM: point estimate missing",
    "Delta lnM: sampling variance missing",
    "SAFE lnM: point estimate missing",
    "SAFE lnM: sampling variance missing",
    "Rescued by SAFE (point available when Delta missing)",
    "Rescued by SAFE (variance available when Delta missing)"
  ),
  Count = c(
    n_total,
    sum(is.na(dat$yi_lnM_raw)),
    sum(is.na(dat$vi_lnM_raw)),
    sum(is.na(dat$yi_lnM_safe)),
    sum(is.na(dat$vi_lnM_safe)),
    sum(is.na(dat$yi_lnM_raw) & !is.na(dat$yi_lnM_safe)),
    sum(is.na(dat$vi_lnM_raw) & !is.na(dat$vi_lnM_safe))
  )
) %>%
  dplyr::mutate(
    Percent = dplyr::if_else(
      Metric == "Total effect sizes",
      NA_real_,
      100 * Count / n_total
    )
  )

# Headline numbers people care about
n_saved <- sum(!is.na(dat$yi_lnM_safe))
p_saved <- 100 * n_saved / n_total

cat(sprintf(
  "\n\n**SAFE produced usable lnM point estimates for %d/%d effect sizes (%.1f%%).**\n\n",
  n_saved, n_total, p_saved
))

summary_tbl %>%
  kableExtra::kable(
    digits = 1,
    caption = "Data availability for lnM estimates: Delta plug-in vs SAFE (counts and % of all effect sizes).",
    align = c("l", "r", "r")
  ) %>%
  kableExtra::kable_styling(full_width = FALSE)
```

### Meta-analysis model 


```{r}
# restrict data just for birds, mammals, amphibians and fish
dat<-dat[which(dat$taxon.for.plot!="reptilia" & 
                 dat$taxon.for.plot!="mollusca" & 
                 dat$taxon.for.plot!="arthropoda"),]
```


::: {.panel-tabset}

### Absulte SMD

```{r}
# meta-analysis using abs(d)
ma_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(ma_abs)
i2_ml(ma_abs)


# adding overall to orchard plot
main <- mod_results(ma_abs, group = "study")
main <- unclass(main)
main$mod_table$name <- "Overall" 
main$data$moderator <- "Overall"
class(main) <- "orchard"

p7 <- orchard_plot(main,  xlab = "abs(SMD)", 
                   group = "study",
                   trunk.size = 0.6,
                   branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#CC6677") 

p7
```

### lnM

```{r}
# meta-analysis using safe
ma_safe <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(ma_safe)
i2_ml(ma_safe)

# adding overll to orchard plot
main2 <- mod_results(ma_safe, group = "study")
main2 <- unclass(main2)
main2$mod_table$name <- "Overall" 
main2$data$moderator <- "Overall"
class(main2) <- "orchard"


p8 <- orchard_plot(main2,  xlab = "lnM", 
                   group = "study",
                   trunk.size = 0.8,
                   branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 


p8
```

:::

### Meta-regression

::: {.panel-tabset}

### Absulte SMD

```{r}
# meta-regression abs(d) using taxon.for.plot
mod_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  mods = ~ taxon.for.plot - 1,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(mod_abs)
# orchard plot
p9 <- orchard_plot(mod_abs, 
                   mod = "taxon.for.plot", 
                   xlab = "abs(SMD)", 
                   group = "study",
                   trunk.size = 0.6,
                   branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = rep("#CC6677", 4))

# multiple comparison
summary(glht(mod_abs, linfct=cbind(contrMat(rep(1,4), type="Tukey"))), test=adjusted("none"))

p9

```

### lnM

```{r}
# meta-regression safe using taxon.for.plot
mod_safe <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ taxon.for.plot -1,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(mod_safe)
# orchard plot
p10 <- orchard_plot(mod_safe, 
                    mod = "taxon.for.plot", 
                    xlab = "lnM", 
                    group = "study",
                    trunk.size = 0.6,
                    branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = rep("#117733", 4)) 
# multiple comparison
summary(glht(mod_safe, linfct=cbind(contrMat(rep(1,4), type="Tukey"))), test=adjusted("none"))

p10
```

:::

### Egger regression (publication bias test)

```{r}
# Creaing a variable using n0
dat$n0 <- (dat$sample.size.control * dat$sample.size.noise.1
                 ) / (dat$sample.size.control + dat$sample.size.noise.1)
dat$nSE <- 1 / sqrt(dat$n0)
dat$nV <- 1 / dat$n0
```


::: {.panel-tabset}

### Absulte SMD

```{r}
# using abs(d)
egger_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  mods = ~ nSE,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(egger_abs)
# bubble plot
p11 <- bubble_plot(egger_abs, mod = "nSE", 
                   ylab = "abs(SMD)", 
                   xlab = "sqrt(inverse sample size)", 
                   group = "study") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#CC6677") 

p11
```

### lnM (sqrt(1/n0))

```{r}
# using safe
egger_safe <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ nSE,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(egger_safe)
# bubble plot
p12 <- bubble_plot(egger_safe, mod = "nSE", 
                   ylab = "lnM", 
                   xlab = "sqrt(inverse sample size)", 
                   group = "study",
                   legend.pos = "bottom.right") +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 

p12
```

### lnM (1/n0)

```{r}
# equivalent to fitting V rather than SE
egger_safe2 <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ nV,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(egger_safe2)
# bubble plot
bubble_plot(egger_safe2, mod = "nV", ylab = "lnM", xlab = "sqrt(inverse sample size)", group = "study")
```


:::


### Sensitivity analysis

```{r}
# sensitivity analysis - we want to get effect sizes where n1 + n2 = or > 40
dat_sens <- dat %>% filter((sample.size.control + sample.size.noise.1) >= 40)
```


::: {.panel-tabset}

### lnM: meta-analysis

```{r}
ma_safe_sens <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat_sens,
  method = "REML",
  test = "t")

summary(ma_safe_sens)
summary(ma_safe)

i2_ml(ma_safe_sens)

orchard_plot(ma_safe_sens,  xlab = "lnM", group = "study")
```

### lnM meta-regression

```{r}
# doing meta-regression again
# meta-regression safe using taxon.for.plot
mod_safe2 <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ taxon.for.plot - 1 + nSE,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat,
  method = "REML",
  test = "t")
summary(mod_safe2)
# comparing
summary(mod_safe)

r2_ml(mod_safe2)

results <- mod_results(mod_safe2, 
                       group = "study", 
                       mod = "taxon.for.plot", 
                       at = list(nSE = c(0)))

# orchard plot
# probably closer to the truth
orchard_plot(results, 
                    mod = "taxon.for.plot", 
                    xlab = "lnM", 
                    group = "study",
                    trunk.size = 0.6,
                    branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = rep("#117733", 4)) 
```


### lnM Egger regression

```{r}
# sensivity + nSE - even smaller

mod_safe3 <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ taxon.for.plot - 1 + nSE,
  random =  list(
    ~ 1 | study,
    ~ 1 |case.nr
  ),
  data   = dat_sens,
  method = "REML",
  test = "t")
summary(mod_safe3)


r2_ml(mod_safe3)

results2 <- mod_results(mod_safe3, 
                       group = "study", 
                       mod = "taxon.for.plot", 
                       at = list(nSE = c(0)))

# orchard plot
# probably closer to the truth
orchard_plot(results2, 
             mod = "taxon.for.plot", 
             xlab = "lnM", 
             group = "study",
             trunk.size = 0.6,
             branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = rep("#117733", 4)) 

```


:::

### Figure 4

```{r}
# making 2 x 3 panel fig
(p7 + p8) / (p9 + p10) / (p11 + p12) + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 16, face = "bold"))
```

# Part II — Simulation reproduction (figures + tables)

## Figures

### Load summary results

```{r}
#| label: sim-load

# to do

results <- readRDS( here::here("results", "lnM_summary_SAFEfun_2025-12-21.rds"))

results <- results %>%
    mutate(facet_label = paste0(design, " n1=", n1, " n2=", n2))

facet_info <- results %>% distinct(facet_label, design, n1, n2)

balanced_ind <- facet_info %>%
    filter(design == "indep", n1 == n2) %>% arrange(n1) %>% pull(facet_label)
unbalanced_ind <- facet_info %>%
    filter(design == "indep", n1 != n2) %>% arrange(n1) %>% pull(facet_label)
paired_balanced <- facet_info %>%
    filter(design == "paired") %>% arrange(n1) %>% pull(facet_label)

new_levels <- c(balanced_ind, unbalanced_ind, paired_balanced)
results <- results %>% mutate(facet_label = factor(facet_label, levels = new_levels))

strip_labeller <- function(x) gsub(" n1=", "   n1=", gsub(" n2=", "   n2=", x))

base_theme <- theme_bw(11) +
    theme(
      legend.title = element_text(),
      legend.position = "right",
      strip.background = element_rect(fill = "grey85", colour = "grey40"),
      strip.text = element_text(face = "plain"),
      panel.grid.minor = element_blank()
)
```

### Figure 1 — Bias of point estimates

```{r}
#| label: fig1-bias
#| fig-cap: "Figure 1. Point-estimate bias (estimate − true lnM). Dashed line indicates 0 bias."

bias_df <- dplyr::bind_rows(
    results %>% dplyr::select(theta, facet_label, delta_bias) %>%
      dplyr::rename(bias = delta_bias) %>% dplyr::mutate(estimator = "PI"),
    results %>% dplyr::select(theta, facet_label, safe_bias) %>%
      dplyr::rename(bias = safe_bias) %>% dplyr::mutate(estimator = "SAFE")
  )

  ggplot(bias_df, aes(theta, bias, colour = estimator, group = estimator)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_line(linewidth = 0.6) +
    facet_wrap(~ facet_label, ncol = 4, nrow = 3,
               labeller = labeller(facet_label = strip_labeller)) +
    labs(x = expression(theta),
         y = "Bias (estimate − true lnM)",
         colour = "estimator") +
    base_theme + 
  theme(legend.position = "bottom")#+
   # annotate("text", x = 0.7, y = max(bias_df$bias, na.rm = TRUE),
   #          label = "Target: 0 bias\nBias typically shrinks\nas θ and n increase",
   #          hjust = 0, vjust = 1, size = 3, colour = "grey25")
```

### Figure 2 — Relative bias of variance estimates

```{r}
#| label: fig2-relbias-var
#| fig-cap: "Figure 2. Relative bias of variance estimates (%). Dashed line indicates 0% (unbiased)."



rb_df <- dplyr::bind_rows(
    results %>% dplyr::select(theta, facet_label, relbias_delta) %>%
      dplyr::rename(relbias = relbias_delta) %>% dplyr::mutate(estimator = "Delta"),
    results %>% dplyr::select(theta, facet_label, relbias_safe) %>%
      dplyr::rename(relbias = relbias_safe)  %>% dplyr::mutate(estimator = "SAFE")
  )

  ggplot(rb_df, aes(theta, relbias, colour = estimator, group = estimator)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_line(linewidth = 0.6) +
    facet_wrap(~ facet_label, ncol = 4, nrow = 3,
               labeller = labeller(facet_label = strip_labeller)) +
    labs(x = expression(theta),
         y = "Relative bias of Var (%)",
         colour = "estimator") +
    base_theme + 
  theme(legend.position = "bottom")#+
    #annotate("text", x = 0.7, y = max(rb_df$relbias, na.rm = TRUE),
    #         label = "+: Var too large (conservative)\n−: Var too small (anti-conservative)",
    #         hjust = 0, vjust = 1, size = 3, colour = "grey25")
```

### Figure S1 — Coverage of nominal 95% intervals

```{r}
#| label: figS1-coverage
#| fig-cap: "Figure S1. Empirical coverage of nominal 95% intervals. Dashed line indicates 0.95."

cov_df <- dplyr::bind_rows(
    results %>% transmute(theta, facet_label, estimator = "PI",   cover = cover_delta),
    results %>% transmute(theta, facet_label, estimator = "SAFE", cover = cover_safe)
  )

  ggplot(cov_df, aes(theta, cover, colour = estimator, group = estimator)) +
    geom_hline(yintercept = 0.95, linetype = "dashed", colour = "grey50") +
    geom_line(linewidth = 0.6) +
    facet_wrap(~ facet_label, ncol = 4, nrow = 3,
               labeller = labeller(facet_label = strip_labeller)) +
    labs(x = expression(theta),
         y = "Empirical coverage",
         colour = "estimator") +
    base_theme + 
  theme(legend.position = "bottom")#+
    #annotate("text", x = 0.7, y = 0.99,
    #         label = "Target: 0.95\nBelow = under-coverage\nAbove = conservative",
    #        hjust = 0, vjust = 1, size = 3, colour = "grey25")

```

### Figure S2 — RMSE of point estimates

```{r}
#| label: figS2-rmse
#| fig-cap: "Figure S2. RMSE of point estimates. Lower values indicate better overall accuracy."


rmse_df <- dplyr::bind_rows(
    results %>% transmute(theta, facet_label, estimator = "PI",   rmse = rmse_delta),
    results %>% transmute(theta, facet_label, estimator = "SAFE", rmse = rmse_safe)
  )

  ggplot(rmse_df, aes(theta, rmse, colour = estimator, group = estimator)) +
    geom_line(linewidth = 0.6) +
    facet_wrap(~ facet_label, ncol = 4, nrow = 3,
               labeller = labeller(facet_label = strip_labeller)) +
    labs(x = expression(theta),
         y = expression(RMSE~"("~hat(lnM)~"\u2212"~lnM~")"),
         colour = "estimator") +
    base_theme + 
  theme(legend.position = "bottom")#+
   #annotate("text", x = 0.7, y = max(rmse_df$rmse, na.rm = TRUE),
    #         label = "RMSE combines\nbias + variance",
    #         hjust = 0, vjust = 1, size = 3, colour = "grey25")
```

### Reading in data for Monte Carlo (MC) standard error (SE)

```{r}
raw_dir <- here("raw_runs")
raw_files <- list.files(raw_dir, pattern = "^row_[0-9]{3}\\.rds$", full.names = TRUE)
if (!length(raw_files)) stop("No raw files found in: ", raw_dir)

row_index_from_path <- function(p) as.integer(sub("^row_([0-9]{3})\\.rds$", "\\1", basename(p)))

# Compute diagnostics for a single grid cell (single raw_df)
one_raw_diag <- function(raw_df) {
  
  # Defensive: remove rep id if present (rep is not used in computations)
  if ("rep" %in% names(raw_df)) raw_df <- raw_df %>% dplyr::select(-rep)
  
  # Match the “ok_d” logic used in the simulation summary:
  # keep only replicates where Delta-method pt & var are defined and var>0
  ok_d <- is.finite(raw_df$delta_pt) & is.finite(raw_df$delta_var) & (raw_df$delta_var > 0)
  
  # If nothing survives the ok_d filter, return NAs
  if (!any(ok_d)) {
    return(data.frame(
      n_ok_d = 0L,
      n_ok_safe = 0L,
      RSE_VarMC_SAFEpt = NA_real_,
      sd_PI = NA_real_,
      sd_SAFE = NA_real_
    ))
  }
  
  df_ok <- raw_df[ok_d, , drop = FALSE]
  
  # Within the ok_d subset, check how many finite SAFE-BC point estimates exist
  ok_safe_pt <- is.finite(df_ok$safe_pt)
  n_ok_safe <- sum(ok_safe_pt)
  
  # Relative standard error (RSE) of the MC variance estimate of SAFE-BC point:
  #   RSE(Var_hat) approx sqrt(2/(n-1)) for sample variance under normality
  RSE_varmc <- if (n_ok_safe >= 3) sqrt(2 / (n_ok_safe - 1)) else NA_real_
  
  # SD of point estimates (needed for MCSE(bias) below)
  # Since true_lnM is constant within a grid cell, sd(est - true) = sd(est).
  sd_PI   <- sd(df_ok$delta_pt, na.rm = TRUE)
  sd_SAFE <- sd(df_ok$safe_pt,  na.rm = TRUE)
  
  data.frame(
    n_ok_d = nrow(df_ok),
    n_ok_safe = n_ok_safe,
    RSE_VarMC_SAFEpt = RSE_varmc,
    sd_PI = sd_PI,
    sd_SAFE = sd_SAFE
  )
}

diag_list <- pbapply::pblapply(raw_files, function(f) {
  rid <- row_index_from_path(f)
  raw_df <- readRDS(f)
  d <- one_raw_diag(raw_df)
  d$row_id <- rid
  d
})

diag_df <- bind_rows(diag_list)

results2 <- results %>%
  mutate(row_id = seq_len(nrow(results))) %>%
  left_join(diag_df, by = "row_id")

```


### Figure S3 — MCSE of bias from raw_runs

```{r}
#| label: figS3-mcse-bias
#| fig-cap: "Figure S3. MCSE of bias (per grid cell) reconstructed from replicate-level files in raw_runs."

# Recompute MCSE(bias) “:
#   MCSE(bias) ≈ sd(estimator)/sqrt(n_eff)
# where n_eff matches the number of usable replicates in each grid cell.
results2 <- results2 %>%
  mutate(
    mcse_bias_PI_disk   = ifelse(is.finite(sd_PI)   & n_ok_d >= 2,    sd_PI   / sqrt(n_ok_d),     NA_real_),
    mcse_bias_SAFE_disk = ifelse(is.finite(sd_SAFE) & n_ok_safe >= 2, sd_SAFE / sqrt(n_ok_safe),  NA_real_)
  )

mc_long <- results2 %>%
  transmute(
    theta, facet_label,
    mcse_bias_PI   = mcse_bias_PI_disk,
    mcse_bias_SAFE = mcse_bias_SAFE_disk
  ) %>%
  pivot_longer(
    cols = c(mcse_bias_PI, mcse_bias_SAFE),
    names_to = "metric", values_to = "value"
  )

p_mc <- ggplot(mc_long, aes(theta, value, colour = metric, group = metric)) +
  geom_line() +
  facet_wrap(~ facet_label, ncol = 4) +
  labs(
    x = expression(theta),
    y = "MCSE of bias",
    title = "Monte-Carlo error diagnostics: MCSE of bias (per grid cell; from raw_runs)"
  ) +
  base_theme + 
  theme(legend.position = "bottom")

print(p_mc)
```

### Figure S4 — MCSE of mean variance estimates (from summary)

```{r}
#| label: figS4-mcse-varmean
#| fig-cap: "Figure S4. MCSE of mean variance estimates (from the simulation summary)."

var_mcse_df <- results %>%
  transmute(theta, facet_label,
            PI_var_MCSE   = mcse_varbar_delta,
            SAFE_var_MCSE = mcse_varbar_safe) %>%
  pivot_longer(cols = c(PI_var_MCSE, SAFE_var_MCSE),
               names_to = "estimator", values_to = "mcse") %>%
  mutate(estimator = recode(estimator,
                            PI_var_MCSE   = "PI (delta-var)",
                            SAFE_var_MCSE = "SAFE (safe-var)"))

p_var_mcse <- ggplot(var_mcse_df, aes(theta, mcse, colour = estimator, group = estimator)) +
  geom_line() +
  facet_wrap(~ facet_label, ncol = 4) +
  labs(x = expression(theta),
       y = "MCSE of mean(variance estimate)",
       colour = NULL,
       title = "Monte-Carlo error of variance estimators (MCSE of the mean)") +
  base_theme + 
  theme(legend.position = "bottom")

print(p_var_mcse)
```


## Tables for the supplement

This section creates compact tables you can paste directly into your HTML supplement.

TODO - requires explanations - and what table column means

### Table S1 — Method feasibility and stability 

```{r}
tab1_stability <- results %>%
  group_by(design, n1, n2, facet_label) %>%
  summarise(
    # at which theta we can define lnM given sample sizes
    theta_min = min(theta, na.rm = TRUE),
    theta_max = max(theta, na.rm = TRUE),
    
    # "how often do we fail?"
    delta_fail_median = median(delta_fail_prop, na.rm = TRUE),
    safe_fail_median  = median(safe_fail_prop,  na.rm = TRUE),
    
    # "how often does SAFE return ok status?"
    SAFE_ok_median    = median(SAFE_ok_rate, na.rm = TRUE),
    
    # "how often does Delta variance get capped?"
    delta_cap_median  = median(delta_cap_rate, na.rm = TRUE),
    
    # also show worst-case over theta (useful for identifying boundary regions)
    delta_fail_max = max(delta_fail_prop, na.rm = TRUE),
    safe_fail_max  = max(safe_fail_prop,  na.rm = TRUE),
    delta_cap_max  = max(delta_cap_rate,  na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  arrange(factor(facet_label, levels = levels(results$facet_label))) %>%
  mutate(
    design = as.character(design),
    facet_label = as.character(facet_label)
  )

kable(tab1_stability, digits = 3,
      caption = "Table S1. Feasibility / stability diagnostics summarised across θ within each design × (n1,n2) facet.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))
```

### Table S2 — SAFE bootstrap acceptance (kept/tried) 

```{r}
tab2_accept <- results %>%
  group_by(design, n1, n2, facet_label) %>%
  summarise(
    q0  = quantile(boot_accept_prop, 0.00, na.rm = TRUE),
    q25 = quantile(boot_accept_prop, 0.25, na.rm = TRUE),
    q50 = quantile(boot_accept_prop, 0.50, na.rm = TRUE),
    q75 = quantile(boot_accept_prop, 0.75, na.rm = TRUE),
    q100= quantile(boot_accept_prop, 1.00, na.rm = TRUE),
    mean = mean(boot_accept_prop, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(factor(facet_label, levels = levels(results$facet_label)))

kable(tab2_accept, digits = 4,
      caption = "Table S2. SAFE acceptance rate (kept/tried) summarised across θ for each facet. Quantiles highlight worst-case acceptance.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))
```

### Table S3 — MCSE diagnostics

```{r}

has_mcse_bias <- all(c("mcse_bias_delta","mcse_bias_safe") %in% names(results))

tab3_mcse <- results %>%
  group_by(design, n1, n2, facet_label) %>%
  summarise(
    # bias MCSE (optional)
    mcse_bias_PI_max   = if (has_mcse_bias) max(mcse_bias_delta, na.rm = TRUE) else NA_real_,
    mcse_bias_SAFE_max = if (has_mcse_bias) max(mcse_bias_safe,  na.rm = TRUE) else NA_real_,
    
    # variance MCSE (these are in your results)
    mcse_var_PI_max    = max(mcse_varbar_delta, na.rm = TRUE),
    mcse_var_SAFE_max  = max(mcse_varbar_safe,  na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  arrange(factor(facet_label, levels = levels(results$facet_label)))

kable(tab3_mcse, digits = 5,
      caption = "Table S3. Monte-Carlo precision (max MCSE over θ within each facet).") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped","hover","condensed"))

```


# Session info
```{r}
#| label: session-info
sessionInfo()
```

