---
title: "Supplementary Tutorial: A new effect size for meta-analysis of magnitude: lnM"
subtitle: "Worked examples + extra results from the simulation (figures + tables)"
date: last-modified
date-format: long
author:
  - name: Shinichi Nakagawa
    attributes:
      corresponding: true
  - name: Ayumi Mizuno
  - name: Coralie Williams
  - name: Malgorzata Lagisz
  - name: Rose E. O'Dea
  - name: Daniel W. A. Noble
  - name: Alistair M. Senior
  - name: Erick Lundgren
  - name: Santiago Ortega
format: 
  html:
    toc: true
    toc-location: left
    toc-depth: 3
    fig-dpi: 300
    toc-float: false
    toc-title: "**CONTENTS**"
    output-file: "index.html"
    embed-resources: true
    code-fold: false
    smooth-scroll: true
    anchor-sections: true
    code-tools: 
      source: false
    number-sections: false
    bibliography: ./bib/ref.bib
    fontsize: "12"
    mainfont: "Roboto"
    monofont: "Roboto"
    # fontcolor: ""
    page-layout: full
    code-overflow: wrap
    df_print: paged
    theme: united
    code-line-numbers: false
    # grid:
    #   margin-width: 190px
execute:
  echo: true
  warning: false
  message: false
  cache: false
---

```{=html}
<style>
  /* --- GLOBAL FONTS & COLORS --- */
  :root {
    --theme-orange: #E95420;
    --theme-grey: #f0f0f0;
  }
  
  body {
    background-color: #fafafa; /* Very soft grey background to reduce eye strain */
    color: #2c3e50;
  }

  /* --- HERO TITLE BLOCK --- */
  header#title-block-header {
    background: white;
    padding: 3rem 2rem;
    margin-bottom: 3rem;
    border-bottom: 4px solid var(--theme-orange);
    text-align: center;
    box-shadow: 0 4px 6px -1px rgba(0,0,0,0.1);
  }
  
  h1.title {
    color: var(--theme-orange);
    font-weight: 900;
    letter-spacing: -1px;
    font-size: 2.5rem;
  }
  
  p.subtitle {
    font-size: 1.3rem;
    color: #7f8c8d;
    font-weight: 300;
  }

  /* --- AUTHORS SECTION --- */
  .quarto-title-meta {
    justify-content: center !important; /* Center the authors */
    text-align: center;
  }

  /* --- HEADINGS --- */
  h2 {
    border-bottom: 2px solid var(--theme-grey);
    padding-bottom: 10px;
    margin-top: 3rem;
    color: #333;
    font-weight: 700;
  }
  
  h2::after {
    /* Little orange tab under headers */
    content: '';
    display: block;
    width: 50px;
    height: 4px;
    background: var(--theme-orange);
    margin-top: 10px;
  }

  /* --- TABLE OF CONTENTS (TOC) --- */
  #quarto-toc-target {
    border-left: 1px solid #e0e0e0;
  }
  
  #TOC > ul > li.active > a {
    color: var(--theme-orange) !important;
    font-weight: bold;
    border-left: 3px solid var(--theme-orange);
    padding-left: 10px;
  }

  /* --- CODE BLOCKS --- */
  div.sourceCode {
    border-radius: 8px;
    box-shadow: 0 4px 6px -1px rgba(0,0,0,0.05);
    border: 1px solid #e1e1e1;
  }

  /* --- LINKS --- */
  a {
    color: var(--theme-orange);
    text-decoration: none;
    transition: all 0.2s;
  }
  a:hover {
    color: #bf360c;
    text-decoration: underline;
  }
  /* Keep Quarto tab headers on one row and make tab boundaries visible */
.nav-tabs {
  flex-wrap: nowrap;
  overflow-x: auto;
  white-space: nowrap;
}
.nav-tabs .nav-link {
  border: 1px solid #ddd;
  margin-right: 4px;
  border-radius: 6px 6px 0 0;
}
.nav-tabs .nav-link:not(.active) {
  background: #f7f7f7;
}
.panel-text { min-height: 5em; }
</style>
```

```{r}
#| label: setup
#| include: true
#| code-fold: true
#| message: false
#| warning: false

# ------------------------------------------------------------------------------
# SETUP
# This chunk:
#   1) Sets global chunk options (figures, widths)
#   2) Loads required packages
#   3) Defines helper functions used throughout the supplement
# ------------------------------------------------------------------------------

knitr::opts_chunk$set(
  fig.align = "center",
  fig.width = 10,
  fig.height = 6,
  out.width = "100%"
)

suppressPackageStartupMessages({
  library(tidyverse)
  library(metafor)        # escalc(), rma.mv()
  library(here)           # robust file paths
  library(orchaRd)        # orchard_plot(), bubble_plot(), .safe_lnM_indep()
  library(kableExtra)     # HTML tables
  library(patchwork)      # combining ggplots   
  library(multcomp)        # multiple comparisons
})

# ---- Ensure compatible orchaRd version -------------------------------------
orchard_min_version <- "2.1.3"  # version that includes SAFE lnM + seed argument

if (packageVersion("orchaRd") < package_version(orchard_min_version)) {
  stop(
    "This tutorial requires orchaRd >= ", orchard_min_version, ".\n",
    "Your version is ", packageVersion("orchaRd"), ".\n\n",
    "Please update by running:\n",
    "  install.packages('orchaRd')\n",
    "or, if needed:\n",
    "  remotes::install_github('daniel1noble/orchaRd')"
  )
}

# ---- Check that the required SAFE function is available ---------------------
if (!exists(".safe_lnM_indep", where = asNamespace("orchaRd"), inherits = FALSE)) {
  stop(
    "This tutorial requires orchaRd:::.safe_lnM_indep(), but it was not found.\n",
    "Please update orchaRd to a version that includes SAFE lnM."
  )
}

# ---- SAFE lnM (orchaRd) -----------------------------------------------------
# Note on purrr::pmap_dfr():
# pmap_dfr() applies a function row-by-row to several input columns (m1i, m2i, sd1i, ...),
# returning a data frame with one output row per effect size.
#
# Settings below match the simulation defaults used in this project.
# Passing a fixed seed makes each SAFE estimate reproducible.

get_lnM_safe <- function(m1, m2, s1, s2, n1, n2, 
                         min_kept = 100000, 
                         chunk_init = 5000, 
                         seed=123) {  
  out <- orchaRd:::.safe_lnM_indep(x1bar = m1, x2bar = m2,
  sd1 = s1, sd2 = s2,
  n1 = n1, n2 = n2,
  min_kept = min_kept,
  chunk_init = chunk_init,seed=seed)

tibble::tibble(
  yi_lnM_safe = out$lnM_SAFE,
  vi_lnM_safe = out$var_lnM_SAFE,
  draws_kept  = out$kept,
  draws_total = out$total,
  attempts    = out$attempts,
  status      = out$status
)
}



# ---- Folded-normal helper for magnitude of SMD ------------------------------
# This converts a signed normal estimate (mu, var) into the expected magnitude
# under the implied folded-normal distribution.

folded_norm <- function(mu, var) {
  sigma  <- sqrt(var)
  if (sigma == 0) return(c(point = abs(mu), var = 0))

  z      <- mu / sigma
  meanFN <- sigma * sqrt(2 / pi) * exp(-0.5 * z^2) +
    mu    * (2 * pnorm(z) - 1)
  varFN  <- mu^2 + var - meanFN^2

  c(point = meanFN, var = varFN)
}

# ---- Delta plug-in lnM (analytic) -------------------------------------------
# IMPORTANT: This requires lnM_delta1_indep() to be available, source it here
source(here("R", "SAFE_fun.R")) 

get_lnM_raw <- function(m1, m2, s1, s2, n1, n2) {
  out <- lnM_delta1_indep(m1, m2, s1, s2, n1, n2)   # returns c(point, var)
  tibble(
    yi_lnM_raw = out["point"],
    vi_lnM_raw = out["var"]
  )
}






```

```{=html}
<!--get_lnM_safe <- function(m1, m2, s1, s2, n1, n2,min_kept = 100000, chunk_init = 5000) {
  out <- safe_lnM_indep(m1, m2, s1, s2, n1, n2,min_kept = min_kept,
    chunk_init = chunk_init)     # returns list(point,var,kept,total)
  tibble(
    yi_lnM_safe = out$point,
    vi_lnM_safe = out$var,
    draws_kept  = out$kept,
    draws_total = out$total
  )
}-->
```

# Overview

This supplement has two parts:

**Part I — Worked examples (Examples 1–2).**

We demonstrate a complete, end-to-end workflow for quantifying effect magnitude using lnM. For each dataset, we first compute a conventional standardized mean difference (SMD) and compare results obtained using absolute SMD–based measures with those obtained using lnM.

We note that we estimate lnM with two approaches:

(i) a **Delta plug-in estimator** and
(ii) the **SAFE bootstrap bias-corrected point estimate (SAFE-BC)** which is designed to remain usable when the plug-in estimator is undefined (e.g., when $MS_B \le MS_W$).

Using these estimates, we then fit multilevel meta-analytic models and meta-regressions and visualize results using the same plot types reported in the main manuscript.

**Part II — Simulation reproduction (Figures + Tables).**

We reproduce all simulation-based figures and tables reported in the manuscript using pre-computed simulation summaries and, where required, outputs from individual simulation replicates (i.e. results from each Monte Carlo run within a design cell). These results characterise estimator performance in terms of bias, variance behaviour, empirical coverage, root mean square error (RMSE), and Monte Carlo standard error (MCSE).

We additionally provide compact tables summarising estimator feasibility, numerical stability, and SAFE bootstrap acceptance rates across simulation designs.

# Part I — Worked examples

Part I illustrates how lnM is computed, analysed, and interpreted in real meta-analytic datasets using two published examples with contrasting data structures.

## Example 1 — Almeida et al nutrition dataset

Almeida *et al.* [-@almeida2021factors] examined whether experimental manipulations of nutrition produce consistently large deviations from control conditions across a range of physiological and performance traits. Because the biological interest lies in the magnitude of responses rather than their direction, this dataset provides a natural setting for comparing analyses based on absolute SMD measures with corresponding analyses based on lnM.

### Effect size calculations

This section calculates four effect-size quantities from the same raw inputs (group means, SDs, and sample sizes). Each step adds new columns to `dat`.

```{r}
#| label: load-data example 1

# Load the dataset from CSV files

dat <- read_csv(here("data", "data_almeida.csv"), show_col_types = FALSE)
```

::: callout-note
| Type | Variable Names | Description |
|:-----------------------|:-----------------------|:-----------------------|
| **Inputs** (Group 1) | `m1i`, `sd1i`, `n1i` | Mean, SD, and Sample Size for Group 1 |
| **Inputs** (Group 2) | `m2i`, `sd2i`, `n2i` | Mean, SD, and Sample Size for Group 2 |
| **Output** (1) | `yi_d`, `vi_d` | Signed SMD |
| **Output** (2) | `yi_d_abs`, `vi_d_abs` | Magnitude of SMD via folded-normal |
| **Output** (3) | `yi_lnM_raw`, `vi_lnM_raw` | lnM via Delta plug-in |
| **Output** (4) | `yi_lnM_safe`, `vi_lnM_safe` | lnM via SAFE |
:::

:::::::: columns
:::: {.column width="49%"}
::: panel-text
## Conventional SMD (signed)

We first compute the standardized mean difference (SMD). This is the conventional effect size used in many meta-analyses.\
`metafor::escalc()` returns a *point estimate* (`yi_d`) and its *sampling variance* (`vi_d`).
:::

```{r}
#| label: escalc-smdq
#| message: false
#| warning: false


dat <- escalc(
  measure = "SMD",
  m1i = m1i, m2i = m2i,
  sd1i = sd1i, sd2i = sd2i,
  n1i = n1i, n2i = n2i,
  data = dat, append = TRUE,
  var.names = c("yi_d", "vi_d")
)
```
::::

::: {.column width="2%"}
:::

:::: {.column width="49%"}
::: panel-text
## Folded-normal magnitude of SMD

Many researchers summarise “effect magnitude” by taking $|SMD|$. However, absolute values are biased upward, especially when sampling error is large.\
Instead of taking $|SMD|$ directly, we use the folded-normal distribution implied by $(\hat{d}, \mathrm{Var}(\hat{d}))$, yielding `yi_d_abs` and `vi_d_abs`.
:::

```{r}
#| label: folded-normal-smdq
#| message: false
#| warning: false

dat <- dat %>%
  mutate(
    abs_d = map2_dfr(yi_d, vi_d, ~{
      out <- folded_norm(.x, .y)
      tibble(yi_d_abs = out["point"], vi_d_abs = out["var"])
    })
  ) %>%
  unnest(abs_d)
```
::::
::::::::

::: callout-note
**About `pmap_dfr()` (row-wise function application)**\
`purrr::pmap_dfr()` applies a function once per row using multiple columns as inputs (e.g., means, SDs, and sample sizes), returning a data frame with one output row per effect size.
:::

:::::::: columns
:::: {.column width="49%"}
::: panel-text
## lnM via Delta plug-in estimator

We compute lnM using an analytic (Delta-method) plug-in estimator. This can be undefined near boundary conditions, which we diagnose later. The result is stored as `yi_lnM_raw` with sampling variance `vi_lnM_raw`.
:::

```{r}
#| label: delta-lnMq
#| message: false
#| warning: false

dat <- dat %>%
  mutate(
    lnM_raw = purrr::pmap_dfr(
      list(m1i, m2i, sd1i, sd2i, n1i, n2i),
      get_lnM_raw
    )
  ) %>%
  unnest(lnM_raw)

```
::::

::: {.column width="2%"}
:::

:::: {.column width="49%"}
::: panel-text
## lnM via SAFE

We compute lnM using SAFE (single-fit parametric bootstrap), which is designed to remain usable when the plug-in estimator fails. The result is stored as `yi_lnM_safe` with sampling variance `vi_lnM_safe`.
:::

```{r}
#| label: SAFE-lnMq
#| message: false
#| warning: false


dat <- dat %>%
  dplyr::mutate(
    lnM_safe = purrr::pmap_dfr(
      list(m1i, m2i, sd1i, sd2i, n1i, n2i),
      get_lnM_safe
    )
  ) %>%
  tidyr::unnest(lnM_safe)


```
::::
::::::::

### How many lnM estimates can be obtained?

The Delta plug-in estimator of lnM can be undefined for some effect sizes, particularly near boundary conditions. SAFE is designed to be more robust in these cases by relying on a single-fit parametric bootstrap. The table below summarises (i) how often each method yields a usable lnM point estimate and sampling variance, and (ii) how often SAFE provides valid estimates in cases where the Delta plug-in estimator fails.

::: callout-note
Percentages are computed relative to the total number of effect sizes in the dataset.
:::

Because lnM can be undefined when the between-group variance is not estimatable, we first quantify how often each estimator returns usable lnM point estimates and sampling variances.

```{r}
#| label: lnM-availability-summary
#| message: false
#| warning: false
#| code-fold: true

stopifnot(exists("dat"), is.data.frame(dat))

n_total <- nrow(dat)

# ---- Helper counts -----------------------------------------------------------

summary_tbl <- tibble::tibble(
  Metric = c(
    "Total effect sizes",
    "Delta plug-in lnM: point estimate unavailable (undefined)",
    "Delta plug-in lnM: sampling variance unavailable",
    "SAFE lnM: point estimate unavailable",
    "SAFE lnM: sampling variance unavailable",
    "SAFE rescue: point estimate available when Delta plug-in is undefined",
    "SAFE rescue: sampling variance available when Delta plug-in is undefined"
  ),
  Count = c(
    n_total,
    sum(is.na(dat$yi_lnM_raw)),
    sum(is.na(dat$vi_lnM_raw)),
    sum(is.na(dat$yi_lnM_safe)),
    sum(is.na(dat$vi_lnM_safe)),
    sum(is.na(dat$yi_lnM_raw) & !is.na(dat$yi_lnM_safe)),
    sum(is.na(dat$vi_lnM_raw) & !is.na(dat$vi_lnM_safe))
  )
) %>%
  dplyr::mutate(
    Percent = dplyr::if_else(
      Metric == "Total effect sizes",
      NA_real_,
      100 * Count / n_total
    )
  )

# Headline numbers people care about
n_saved <- sum(!is.na(dat$yi_lnM_safe))
p_saved <- 100 * n_saved / n_total

cat(sprintf(
  "\n\n**SAFE produced defined lnM point estimates for %d/%d effect sizes (%.1f%%).**\n\n",
  n_saved, n_total, p_saved
))

summary_tbl %>%
  kableExtra::kable(
    digits = 1,
    caption = "Data availability for lnM estimates: Delta plug-in vs SAFE (counts and % of all effect sizes).",
    align = c("l", "r", "r")
  ) %>%
  kableExtra::kable_styling(full_width = FALSE)%>%
  kableExtra::scroll_box(width = "100%", height = "400px")
```

::: callout-tip
Rows labelled “SAFE rescue” count effect sizes for which the Delta plug-in estimator failed but SAFE returned a valid estimate.
:::

This table shows that the Delta plug-in estimator fails to return a defined lnM point estimate or sampling variance for a substantial fraction of effect sizes in this dataset. In contrast, SAFE returns both a point estimate and a sampling variance for all effect sizes. As a result, SAFE “rescues” a large number of comparisons that would otherwise need to be excluded from lnM-based analyses.

### Meta-analysis model

In this worked example we focus on **body condition** outcomes from the Almeida dataset. Almeida *et al.* report multiple outcome categories (e.g., condition and other trait classes); to keep the tutorial concrete and comparable across methods, we restrict to the `Category == "Condition"` subset.

We then fit the same multilevel random-effects structure for two magnitude metrics: (i) the folded-normal magnitude of SMD and (ii) lnM estimated with SAFE. In both cases, effect sizes are nested within studies (random intercepts for `Study` and for the within-study effect-size ID, `ES.ID`).

```{r}
#| label: subset-condition-example-1


dat_cond <- dat %>% filter(Category == "Condition") %>%
  rename(exposure = magnitude)
dim(dat_cond)
```

:::::: columns
::: {.column width="49%"}
## Absolute SMD (folded-normal)
::: panel-text
We meta-analyse the SMD magnitude using the folded-normal transformation.
:::
```{r}
#| label: ma-abs-smd
#| message: false
#| warning: false

ma_abs <- rma.mv(
  yi     = yi_d_abs,
  V      = vi_d_abs,
  random = list(~ 1 | Study, ~ 1 | ES.ID),
  data   = dat_cond,
  method = "REML",test   = "t"
)

summary(ma_abs)
i2_ml(ma_abs)
ml_m1(ma_abs)


main <- orchaRd::mod_results(ma_abs, group = "Study")
main <- unclass(main)
main$mod_table$name <- "Overall"
main$data$moderator <- "Overall"
class(main) <- "orchard"

p1 <- orchard_plot(main,  xlab = "abs(SMD)", 
                   group = "Study",
                   trunk.size = 0.6,
                   branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#CC6677") 

p1
```
:::

::: {.column width="2%"}
:::

::: {.column width="49%"}
## lnM (SAFE)
::: panel-text
We meta-analyse lnM using SAFE estimates.
:::
```{r}
#| label: ma-lnM-safe
#| message: false
#| warning: false

ma_safe <- rma.mv(
  yi     = yi_lnM_safe,
  V      = vi_lnM_safe,
  random = list(~ 1 | Study, ~ 1 | ES.ID),
  data   = dat_cond,
  method = "REML", test   = "t"
)

summary(ma_safe)
i2_ml(ma_safe)
ml_m1(ma_safe)


main2 <- orchaRd::mod_results(ma_safe, group = "Study")
main2 <- unclass(main2)
main2$mod_table$name <- "Overall"
main2$data$moderator <- "Overall"
class(main2) <- "orchard"

p2 <- orchard_plot(main2,  xlab = "lnM", 
                   group = "Study",
                   trunk.size = 0.6,
                   branch.size = 5) +
  scale_colour_manual(values = rep("grey20",8)) +
  scale_fill_manual(values = "#117733") 

p2

```
:::
::::::

::: callout-note
## Interpreting lnM on an SMD scale

For interpretability, we map $lnM$ onto an SMD-equivalent effect size, $d_{\mathrm{eq}}$, which is expressed in units of the within-group standard deviation. Using the large-sample approximation:

$d_{\mathrm{eq}}  \approx \sqrt{2} e^{\ln M}$

```{r}
#| label: lnm-to-deq-approx
#| message: false
#| warning: false
#| code-fold: true

lnm_to_deq_approx <- function(model, label = "model") {
  b0_hat <- as.numeric(coef(model)[1])
  se_b0  <- sqrt(vcov(model)[1, 1])

  df_b0 <- tryCatch(as.numeric(model$ddf[1]), error = function(e) NA_real_)
  crit  <- if (is.finite(df_b0)) qt(0.975, df = df_b0) else qnorm(0.975)

  ci_lb <- b0_hat - crit * se_b0
  ci_ub <- b0_hat + crit * se_b0

  deq_hat <- sqrt(2) * exp(b0_hat)
  deq_ci  <- sqrt(2) * exp(c(ci_lb, ci_ub))

  cat(sprintf(
    "%s: d_eq (approx) = %.2f (95%% CI: %.2f to %.2f)\n",
    label, deq_hat, deq_ci[1], deq_ci[2]
  ))
}

lnm_to_deq_approx(ma_safe, label = "Overall model (ma_safe)")
```

Because this transformation is monotone (larger lnM always implies larger $d_{\mathrm{eq}}$), we can obtain a 95% interval for $d_{\mathrm{eq}}$ by transforming the lower and upper endpoints of the lnM interval.
:::
In the Almeida dataset, the folded-normal SMD magnitude is constrained to be non-negative and many estimates accumulate near zero, with a small number of extreme values extending the upper tail. lnM, in contrast, is defined on an unbounded scale and spreads effect magnitudes more evenly, making between-study differences in overall magnitude easier to resolve when direction is not of primary interest.

### Meta-regression

We next examine whether variation in effect magnitude is associated with study-level covariates. Specifically, we regress each magnitude metric on exposure intensity, exposure duration, and recovery period while retaining the same multilevel structure (random intercepts for `Study` and within-study effect-size ID, `ES.ID`).

:::::: columns
::: {.column width="49%"}
## Absolute SMD (folded-normal)

```{r}
#| label: mod-abs-smd
#| message: false
#| warning: false

mod_abs <- rma.mv(
  yi   = yi_d_abs,
  V    = vi_d_abs,
  mods = ~ exposure + duration + Recovery,
  random = list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML", test   = "t"
  )

summary(mod_abs)

p3 <- bubble_plot(
  mod_abs, mod = "Recovery",
  ylab = "abs(SMD)",
  xlab = "Recovery period",
  group = "Study"
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#CC6677")

p3
```
:::

::: {.column width="2%"}
:::

::: {.column width="49%"}
## lnM (SAFE)

```{r}
#| label: mod-lnM-safe
#| message: false
#| warning: false

mod_safe <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ exposure + duration + Recovery,
  random = list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML", test   = "t"
)

summary(mod_safe)

p4 <- bubble_plot(
  mod_safe, mod = "Recovery",
  ylab = "lnM",
  xlab = "Recovery period",
  group = "Study", legend.pos = "bottom.left"
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#117733")

p4
```
:::
::::::

The folded-normal SMD magnitude is bounded at zero and shows strong compression near the lower end of the scale, which can attenuate moderator slopes. lnM is unbounded and approximately normal, providing greater dynamic range and making moderator relationships (e.g., with recovery period) easier to detect and interpret under the same multilevel structure.

### Small-study effects diagnostic

We use an Egger-style meta-regression [@egger1997bias] to assess whether estimated effect magnitudes systematically vary with study size, a pattern that can be consistent with small-study effects (including publication bias). In this approach, effect sizes are regressed on a measure of study precision rather than being compared visually in a funnel plot.

For lnM, using the usual sampling standard error as a predictor is problematic because lnM and its estimated standard error are both functions of the same underlying variance components. This creates a structural dependence between the response and predictor that violates the assumptions of Egger-type regressions. Following recent recommendations for multilevel meta-analyses, we therefore use a *sample-size-based* measure of precision instead [@nakagawa2022methods].

We define the harmonic-mean (or “effective”) sample size as:

$$
n_0 = \frac{n_1 n_2}{n_1 + n_2}
$$

and fit two related models using:

-   **1 / sqrt(n₀)**, an *SE-analogue*, to test for the presence of small-study effects; and\
-   **1 / n₀**, a *variance-analogue*, to obtain a bias-adjusted overall effect estimate if evidence of small-study effects is detected.

Both predictors depend only on sample sizes and are therefore independent of the observed lnM values themselves. This allows small-study effects to be assessed without inducing artefactual correlations between effect sizes and their precision.

### Worked example: Egger-style models using n₀

```{r}
#| label: ex1-egger-setup
#| message: false
#| warning: false
# Create sample-size-based predictors using the harmonic mean sample size n0
dat_cond$n0 <- (dat_cond$n1i * dat_cond$n2i) / (dat_cond$n1i + dat_cond$n2i)
dat_cond$nSE <- 1 / sqrt(dat_cond$n0)
dat_cond$nV <- 1 / dat_cond$n0
```

::: callout-note
### How to read these models

The models using **1 / sqrt(n₀)** test for the presence of small-study effects (Egger-style slope test).\
The model using **1 / n₀** provides a bias-adjusted estimate of the overall effect size, evaluated at infinite sample size.
:::

:::::: columns
::: {.column width="49%"}
## Absolute SMD (folded-normal; 1/ sqrt(n₀))

```{r}
#| label: ex1-egger-abs-smd
#| message: false
#| warning: false

egger_abs <- rma.mv(
  yi   = yi_d_abs, 
  V    = vi_d_abs,
  mods = ~ nSE,
  random = list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test   = "t"
)

summary(egger_abs)

p5 <- bubble_plot(
  egger_abs, mod = "nSE",
  ylab = "abs(SMD)",
  xlab = "sqrt(inverse sample size)",
  group = "Study"
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#CC6677")

p5
```
:::

::: {.column width="2%"}
:::

::: {.column width="49%"}
## lnM (SAFE; 1 / sqrt(n₀))

```{r}
#| label: ex1-egger-lnM-safe
#| message: false
#| warning: false

egger_safe <- rma.mv(
  yi   = yi_lnM_safe, 
  V    = vi_lnM_safe,
  mods = ~ nSE,
  random = list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test   = "t"
)

summary(egger_safe)

p6 <- bubble_plot(
  egger_safe, mod = "nSE",
  ylab = "lnM",
  xlab = "sqrt(inverse sample size)",
  group = "Study",legend.pos = "bottom.right"
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#117733")

p6
```
:::
::::::
For the folded-normal absolute SMD, the fitted slope is shallow and the bulk of estimates cluster near the lower bound of the scale, indicating limited sensitivity of absolute SMD to study size and little ability to distinguish small-study effects. In contrast, lnM exhibits a clearer positive association with 1 / sqrt(n₀), with smaller effective sample sizes corresponding to larger estimated magnitudes. This pattern indicates the presence of small-study effects for lnM and illustrates that lnM provides a wider dynamic range for diagnosing size-related structure in effect magnitudes under the same multilevel model.
::::::: columns
::: {.column width="25%"}
:::

:::: {.column width="50%"}
## lnM (SAFE; 1 / n₀ — bias-adjusted estimate)

```{r}

#| label: ex1-egger-lnM-safe-2
#| message: false
#| warning: false

# Variance-analogue specification for bias adjustment
egger_safe2 <- rma.mv(
  yi   = yi_lnM_safe,
  V    = vi_lnM_safe,
  mods = ~ nV,
  random = list(
    ~ 1 | Study,
    ~ 1 | ES.ID
  ),
  data   = dat_cond,
  method = "REML",
  test   = "t"
)

summary(egger_safe2)

bubble_plot(
  egger_safe2, mod = "nV",
  ylab = "lnM",
  xlab = "inverse sample size",
  group = "Study"
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#117733")

```

::: callout-note
## Mapping lnM to an SMD-equivalent effect size

For interpretability, we convert the intercept from the **1 / n₀** model (the bias-adjusted lnM) to an SMD-equivalent magnitude using the large-sample approximation.

```{r}
#| label: lnm-to-deq-approx-egger
lnm_to_deq_approx(egger_safe2, "Egger-style model (egger_safe2)")
```
:::
::::

::: {.column width="25%"}
:::
:::::::

### Sensitivity analysis

We repeated the lnM analyses after restricting effect sizes to comparisons with total sample size **n₁ + n₂ ≥ 40**.

```{r}
#| label: ex1-sensitivity-subset
#| message: false
#| warning: false

dat_cond_sens <- dat_cond %>% dplyr::filter((n1i + n2i) >= 40)
dim(dat_cond_sens)
```

### lnM: meta-analysis (sensitivity subset)

```{r}
#| label: ma-lnM-safe-sens
#| message: false
#| warning: false

ma_safe_sens <- rma.mv(
  yi     = yi_lnM_safe, 
  V      = vi_lnM_safe,
  random = list(~ 1 | Study, ~ 1 | ES.ID),
  data   = dat_cond_sens,
  method = "REML",
  test   = "t"
)

summary(ma_safe_sens)
i2_ml(ma_safe_sens)

orchard_plot(ma_safe_sens, xlab = "lnM", group = "Study")
```

::: callout-note
## Mapping lnM to an SMD-equivalent effect size

```{r}
#| label: lnm-to-deq-approx-sens
#| message: false
#| warning: false

lnm_to_deq_approx(ma_safe_sens, "Sensitivity model (n₁ + n₂ ≥ 40)")
```
:::

#### lnM: meta-regression (main vs sensitivity)

To compare moderator patterns, we fit the same meta-regression model to the full dataset and to the sensitivity subset.

:::::: columns
::: {.column width="49%"}
## Main dataset

```{r}
#| label: mod-lnM-safe-main
#| message: false
#| warning: false

mod_safe_main <- rma.mv(
  yi     = yi_lnM_safe, 
  V      = vi_lnM_safe,
  mods   = ~ exposure + duration + Recovery + nSE,
  random = list(~ 1 | Study, ~ 1 | ES.ID),
  data   = dat_cond,
  method = "REML",
  test   = "t"
)

summary(mod_safe_main)
r2_ml(mod_safe_main)

bubble_plot(
  mod_safe_main, mod = "Recovery",
  ylab = "lnM",
  xlab = "Recovery period",
  group = "Study",
  legend.pos = "bottom.left"
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#117733")
```
:::

::: {.column width="2%"}
:::

::: {.column width="49%"}
## Sensitivity subset (n₁ + n₂ ≥ 40)

```{r}
#| label: mod-lnM-safe-sens
#| message: false
#| warning: false

mod_safe_sens <- rma.mv(
  yi     = yi_lnM_safe, 
  V      = vi_lnM_safe,
  mods   = ~ exposure + duration + Recovery + nSE,
  random = list(~ 1 | Study, ~ 1 | ES.ID),
  data   = dat_cond_sens,
  method = "REML",
  test   = "t"
)

summary(mod_safe_sens)
r2_ml(mod_safe_sens)

bubble_plot(
  mod_safe_sens, mod = "Recovery",
  ylab = "lnM",
  xlab = "Recovery period",
  group = "Study",
  legend.pos = "bottom.left"
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#117733")

```
:::
::::::

These sensitivity results indicate that lnM-based inferences are robust to the exclusion of small-sample comparisons.

### Figure 3

Figure 3 collates the three analyses above (overall model, meta-regression, and small-study effects diagnostic) to highlight how inference differs when magnitude is quantified using folded-normal SMD versus lnM (SAFE).

```{r}
#| label: ex1-figure-3
# Combine panels into a 2 × 3 layout:
# Row 1: overall meta-analyses (absolute SMD vs lnM)
# Row 2: meta-regressions (absolute SMD vs lnM)
# Row 3: Egger-style diagnostics (absolute SMD vs lnM)
(p1 + ggtitle("SMD estimates") + p2 + ggtitle("lnM estimates")) / 
  (p3 + p4) / 
  (p5 + p6) + 
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(size = 16, face = "bold"))
```

## Example 2 — Kunc & Schmidt anthropogenic noise dataset

This dataset synthesises experiments testing whether exposure to **anthropogenic noise** alters animal behaviour or physiology relative to control (quiet) conditions [@kunc2019effects]. Each effect size contrasts a noise treatment with a control group. In many applications the direction of change can also be important; here, we focus on **magnitude** to illustrate how lnM can summarise the strength of responses across taxa and response variables.

### Effect size calculations

This section mirrors Example 1. We compute a conventional standardized mean difference (here using the unequal-variance variant, SMDH), its folded-normal magnitude, and lnM using both the Delta plug-in and SAFE.

```{r}
#| label: load-data-example-2
#| message: false
#| warning: false

# Load data
dat_raw <- read.csv(here("data", "kunc_2019.csv"), na.strings = c("", "NA")) %>%
  as.data.frame()

# Basic cleaning: keep rows with species info and usable SDs
dat_raw <- dat_raw %>%
  dplyr::filter(!is.na(species.latin), species.latin != "NA") %>%
  dplyr::mutate(
    sd.control = as.numeric(as.character(sd.control)),
    sd.noise   = as.numeric(as.character(sd.noise))
  ) %>%
  dplyr::filter(!is.na(sd.control), !is.na(sd.noise))

dim(dat_raw)

```

::: callout-note
| Type | Variable Names | Description |
|:-----------------------|:-----------------------|:-----------------------|
| **Inputs (Control)** | `mean.control`, `sd.control`, `sample.size.control` | Mean, SD, and Sample Size in control condition |
| **Inputs (Noise)** | `mean.noise`, `sd.noise`, `sample.size.noise.1` | Mean, SD, and Sample Size in noise condition |
| **Output (1)** | `yi_d`, `vi_d` | Signed SMDH |
| **Output (2)** | `yi_d_abs`, `vi_d_abs` | Magnitude of SMDH via folded-normal |
| **Output (3)** | `yi_lnM_raw`, `vi_lnM_raw` | lnM via Delta plug-in |
| **Output (4)** | `yi_lnM_safe`, `vi_lnM_safe` | lnM via SAFE |
:::

:::::: columns
::: {.column width="49%"}
## Standardized mean difference
::: panel-text
We compute an unequal-variance standardized mean difference (SMDH) using metafor::escalc(). This returns a point estimate (yi_d) and sampling variance (vi_d).
:::
```{r}
#| label: ex2-escalc-smdh
#| message: false
#| warning: false

dat <- escalc(
  measure = "SMDH",
  m1i = mean.control, sd1i = sd.control, n1i = sample.size.control,
  m2i = mean.noise,   sd2i = sd.noise,   n2i = sample.size.noise.1,
  data = dat_raw, append = TRUE,
  var.names = c("yi_d", "vi_d")
) %>%
  dplyr::filter(is.finite(yi_d), is.finite(vi_d))
```
:::

::: {.column width="2%"}
:::

::: {.column width="49%"}
## Folded-normal magnitude of SMDH
::: panel-text
We convert the signed SMDH into a magnitude measure using the folded-normal transformation.
:::
```{r}



#| label: ex2-folded-normal-smd
#| message: false
#| warning: false

dat <- dat %>%
  dplyr::mutate(
    abs_d = purrr::map2_dfr(yi_d, vi_d, ~{
      out <- folded_norm(.x, .y)
      tibble::tibble(yi_d_abs = out["point"], vi_d_abs = out["var"])
    })
  ) %>%
  tidyr::unnest(abs_d)
```
:::
::::::

:::::: columns
::: {.column width="49%"}
## lnM via Delta plug-in estimator
::: panel-text
We compute lnM using the analytic (Delta-method) plug-in estimator. This can be undefined near boundary conditions.
:::
```{r}
#| label: ex2-delta-lnm
#| message: false
#| warning: false

dat <- dat %>%
  dplyr::mutate(
    lnM_raw = purrr::pmap_dfr(
      list(mean.control, mean.noise, sd.control, sd.noise,
           sample.size.control, sample.size.noise.1),
      get_lnM_raw
    )
  ) %>%
  tidyr::unnest(lnM_raw)

```
:::

::: {.column width="2%"}
:::

::: {.column width="49%"}
## lnM via SAFE
::: panel-text
We compute lnM using SAFE (single-fit parametric bootstrap).
:::
```{r}
#| label: ex2-safe-lnm
#| message: false
#| warning: false

dat <- dat %>%
  dplyr::mutate(
    lnM_safe = purrr::pmap_dfr(
      list(mean.control, mean.noise, sd.control, sd.noise,
           sample.size.control, sample.size.noise.1),
      get_lnM_safe
    )
  ) %>%
  tidyr::unnest(lnM_safe)

```
:::
::::::

### How many lnM estimates can be obtained?

```{r}
#| label: lnM-availability-summary-ex2
#| message: false
#| warning: false
#| code-fold: true

stopifnot(exists("dat"), is.data.frame(dat))

n_total <- nrow(dat)

summary_tbl <- tibble::tibble(
  Metric = c(
    "Total effect sizes",
    "Delta plug-in lnM: point estimate unavailable (undefined)",
    "Delta plug-in lnM: sampling variance unavailable",
    "SAFE lnM: point estimate unavailable",
    "SAFE lnM: sampling variance unavailable",
    "SAFE rescue: point estimate available when Delta plug-in lnM is undefined",
    "SAFE rescue: sampling variance available when Delta plug-in lnM is undefined"
  ),
  Count = c(
    n_total,
    sum(is.na(dat$yi_lnM_raw)),
    sum(is.na(dat$vi_lnM_raw)),
    sum(is.na(dat$yi_lnM_safe)),
    sum(is.na(dat$vi_lnM_safe)),
    sum(is.na(dat$yi_lnM_raw) & !is.na(dat$yi_lnM_safe)),
    sum(is.na(dat$vi_lnM_raw) & !is.na(dat$vi_lnM_safe))
  )
) %>%
  dplyr::mutate(
    Percent = dplyr::if_else(Metric == "Total effect sizes", NA_real_, 100 * Count / n_total)
  )

n_saved <- sum(!is.na(dat$yi_lnM_safe))
p_saved <- 100 * n_saved / n_total

cat(sprintf(
  "\n\n**SAFE produced defined lnM point estimates for %d/%d effect sizes (%.1f%%).**\n\n",
  n_saved, n_total, p_saved
))

summary_tbl %>%
  kableExtra::kable(
    digits = 1,
    caption = "Data availability for lnM estimates (Example 2): Delta plug-in vs SAFE (counts and % of all effect sizes).",
    align = c("l", "r", "r")
  ) %>%
  kableExtra::kable_styling(full_width = FALSE) %>%
  kableExtra::scroll_box(width = "100%", height = "400px")
```

### Meta-analysis model

We focus on the main taxa retained in the worked example (birds, mammals, amphibians, and fish), excluding groups with very sparse representation.

```{r}


#| label: ex2-filter-taxa
#| message: false
#| warning: false

dat <- dat %>%
  dplyr::filter(!taxon.for.plot %in% c("reptilia", "mollusca", "arthropoda"))


```

:::::: columns
::: {.column width="49%"}
## Absolute SMD (folded-normal)

```{r}
#| label: ma-abs-smd-ex2
#| message: false
#| warning: false

ma_abs_ex2 <- rma.mv(
  yi     = yi_d_abs,
  V      = vi_d_abs,
  random = list(~ 1 | study, ~ 1 | case.nr),
  data   = dat,
  method = "REML",
  test   = "t"
)

summary(ma_abs_ex2)
i2_ml(ma_abs_ex2)

main <- orchaRd::mod_results(ma_abs_ex2, group = "study")
main <- unclass(main)
main$mod_table$name <- "Overall"
main$data$moderator <- "Overall"
class(main) <- "orchard"

p7 <- orchard_plot(main, xlab = "Folded-normal SMD magnitude", group = "study",
                   trunk.size = 0.6, branch.size = 5) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#CC6677")
p7
```
:::

::: {.column width="2%"}
:::

::: {.column width="49%"}
## lnM (SAFE)

```{r}
#| label: ma-lnM-safe-ex2
#| message: false
#| warning: false

ma_safe_ex2 <- rma.mv(
  yi     = yi_lnM_safe,
  V      = vi_lnM_safe,
  random = list(~ 1 | study, ~ 1 | case.nr),
  data   = dat,
  method = "REML",
  test   = "t"
)

summary(ma_safe_ex2)
i2_ml(ma_safe_ex2)

main2 <- orchaRd::mod_results(ma_safe_ex2, group = "study")
main2 <- unclass(main2)
main2$mod_table$name <- "Overall"
main2$data$moderator <- "Overall"
class(main2) <- "orchard"

p8 <- orchard_plot(main2, xlab = "lnM", group = "study",
                   trunk.size = 0.8, branch.size = 5) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#117733")
p8
```
:::
::::::

::: callout-note
## Mapping lnM to an SMD-equivalent effect size

```{r}
#| label: lnm-to-deq-approx-ex2
#| message: false
#| warning: false

lnm_to_deq_approx(ma_safe_ex2, "Overall model (Example 2)")
```
:::
In this example, absolute SMD estimates again cluster near the lower bound of the scale, with a small number of extreme values dominating visual spread. By contrast, lnM produces a more symmetric and dispersed distribution of magnitudes, allowing differences in response strength across studies to be more clearly resolved. As in Example 1, this illustrates how lnM can summarise the magnitude of treatment effects across heterogeneous outcomes while retaining a scale that is better suited for synthesis and model-based inference.

### Meta-regression

We next examine whether effect magnitudes vary across broad taxonomic groups. Here, `taxon.for.plot` is treated as a categorical moderator (with `- 1` so each group mean is estimated directly).

:::::: columns
::: {.column width="49%"}
## Absolute SMD (folded-normal magnitude)

```{r}
#| label: mod-abs-smd-ex2
#| message: false
#| warning: false

mod_abs_ex2 <- rma.mv(
  yi     = yi_d_abs,
  V      = vi_d_abs,
  mods   = ~ taxon.for.plot - 1,
  random = list(~ 1 | study, ~ 1 | case.nr),
  data   = dat,
  method = "REML",
  test   = "t"
)

summary(mod_abs_ex2)

p9 <- orchard_plot(
  mod_abs_ex2,
  mod = "taxon.for.plot",
  xlab = "Folded-normal SMD magnitude",
  group = "study",
  trunk.size = 0.6,
  branch.size = 5
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = rep("#CC6677", 4))

p9


#Pairwise comparisons among taxon means (Tukey-type)
summary(
  glht(mod_abs_ex2, linfct = cbind(contrMat(rep(1, 4), type = "Tukey"))),
  test = adjusted("none")
)
```
:::

::: {.column width="2%"}
:::

::: {.column width="49%"}
## lnM (SAFE)

```{r}
#| label: mod-lnM-safe-ex2
#| message: false
#| warning: false

mod_safe_ex2 <- rma.mv(
  yi     = yi_lnM_safe,
  V      = vi_lnM_safe,
  mods   = ~ taxon.for.plot - 1,
  random = list(~ 1 | study, ~ 1 | case.nr),
  data   = dat,
  method = "REML",
  test   = "t"
)

summary(mod_safe_ex2)

p10 <- orchard_plot(
  mod_safe_ex2,
  mod = "taxon.for.plot",
  xlab = "lnM (SAFE)",
  group = "study",
  trunk.size = 0.6,
  branch.size = 5
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = rep("#117733", 4))

p10

#Pairwise comparisons among taxon means (Tukey-type)
summary(
  glht(mod_safe_ex2, linfct = cbind(contrMat(rep(1, 4), type = "Tukey"))),
  test = adjusted("none")
)
```
:::
::::::
Using folded-normal SMDs, group means are strongly compressed toward zero and heavily influenced by a small number of extreme values, resulting in substantial overlap among taxa and limited ability to distinguish typical response magnitudes. In contrast, lnM yields a wider and more symmetric distribution of effect magnitudes within groups, revealing clearer separation among taxa and reducing the influence of extreme observations.

### Small-study effects diagnostic

We use the same approach as in Example 1.

```{r}
#| label: ex2-egger-setup
#| message: false
#| warning: false

dat <- dat %>%
  dplyr::mutate(
    n0  = (sample.size.control * sample.size.noise.1) /
          (sample.size.control + sample.size.noise.1),
    nSE = 1 / sqrt(n0),
    nV  = 1 / n0
  )
```

:::::: columns
::: {.column width="49%"}
## Absolute SMD (folded-normal; 1 / sqrt(n₀))

```{r}
#| label: ex2-egger-abs-smd
#| message: false
#| warning: false

egger_abs_ex2 <- rma.mv(
  yi     = yi_d_abs,
  V      = vi_d_abs,
  mods   = ~ nSE,
  random = list(~ 1 | study, ~ 1 | case.nr),
  data   = dat,
  method = "REML",
  test   = "t"
)

summary(egger_abs_ex2)

p11 <- bubble_plot(
  egger_abs_ex2, mod = "nSE",
  ylab = "Folded-normal SMD magnitude",
  xlab = "sqrt(inverse sample size)",
  group = "study"
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#CC6677")

p11
```
:::

::: {.column width="2%"}
:::

::: {.column width="49%"}
## lnM (SAFE; 1 / sqrt(n₀))

```{r}
#| label: ex2-egger-lnM-safe
#| message: false
#| warning: false

egger_safe_ex2 <- rma.mv(
  yi     = yi_lnM_safe,
  V      = vi_lnM_safe,
  mods   = ~ nSE,
  random = list(~ 1 | study, ~ 1 | case.nr),
  data   = dat,
  method = "REML",
  test   = "t"
)

summary(egger_safe_ex2)

p12 <- bubble_plot(
  egger_safe_ex2, mod = "nSE",
  ylab = "lnM (SAFE)",
  xlab = "sqrt(inverse sample size)",
  group = "study",
  legend.pos = "bottom.right"
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#117733")

p12
```
:::
::::::
The folded-normal SMD shows little association with precision, reflecting compression of effect magnitudes near zero. In contrast, lnM exhibits a clear positive relationship with decreasing study size, indicating greater sensitivity to size-dependent differences in effect magnitude.


::::::: columns
::: {.column width="25%"}
:::
::: {.column width="50%"}
## lnM (SAFE; 1 / n₀)
```{r}
#| label: ex2-egger-lnM-safe-2
#| message: false
#| warning: false

egger_safe2_ex2 <- rma.mv(
  yi     = yi_lnM_safe,
  V      = vi_lnM_safe,
  mods   = ~ nV,
  random = list(~ 1 | study, ~ 1 | case.nr),
  data   = dat,
  method = "REML",
  test   = "t"
)

summary(egger_safe2_ex2)

bubble_plot(
  egger_safe2_ex2, mod = "nV",
  ylab = "lnM (SAFE)",
  xlab = "inverse sample size",
  group = "study"
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = "#117733")
```
::: callout-note
## Mapping lnM to an SMD-equivalent effect size
```{r}
#| label: lnm-to-deq-approx-egger-ex2
#| message: false
#| warning: false

lnm_to_deq_approx(egger_safe2_ex2, "Egger-style model (Example 2; 1 / no)")
```
:::
:::
::: {.column width="25%"}
:::
:::::::

### Sensitivity analysis

We repeated the lnM analyses after restricting effect sizes to comparisons with total sample size n₁ + n₂ ≥ 40.

```{r}
#| label: ex2-sensitivity-subset
#| message: false
#| warning: false

dat_sens <- dat %>%
  dplyr::filter((sample.size.control + sample.size.noise.1) >= 40)

dim(dat_sens)
```

## lnM: meta-analysis (sensitivity subset)

```{r}
#| label: ex2-ma-lnM-safe-sens
#| message: false
#| warning: false

ma_safe_sens_ex2 <- rma.mv(
  yi     = yi_lnM_safe,
  V      = vi_lnM_safe,
  random = list(~ 1 | study, ~ 1 | case.nr),
  data   = dat_sens,
  method = "REML",
  test   = "t"
)

summary(ma_safe_sens_ex2)
i2_ml(ma_safe_sens_ex2)

orchard_plot(ma_safe_sens_ex2, xlab = "lnM", group = "study")
```

::: callout-note
## Mapping lnM to an SMD-equivalent effect size

```{r}
#| label: lnm-to-deq-approx-sens-ex2
#| message: false
#| warning: false

lnm_to_deq_approx(ma_safe_sens_ex2, "Sensitivity model (Example 2; n₁ + n₂ ≥ 40)")
```
:::

### lnM: taxon model (main vs sensitivity)

To compare patterns across taxa, we refit the same taxon model in the sensitivity subset.

:::::: columns
::: {.column width="49%"}
## Main dataset

```{r}
#| label: ex2-tax-model-main
#| message: false
#| warning: false

tax_model_main_ex2 <- rma.mv(
  yi     = yi_lnM_safe,
  V      = vi_lnM_safe,
  mods   = ~ taxon.for.plot - 1 + nSE,
  random = list(~ 1 | study, ~ 1 | case.nr),
  data   = dat,
  method = "REML",
  test   = "t"
)

summary(tax_model_main_ex2)
r2_ml(tax_model_main_ex2)

res_main <- mod_results(
  tax_model_main_ex2,
  group = "study",
  mod = "taxon.for.plot",
  at = list(nSE = 0)
)

orchard_plot(
  res_main,
  mod = "taxon.for.plot",
  xlab = "lnM (SAFE)",
  group = "study",
  trunk.size = 0.6,
  branch.size = 5
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = rep("#117733", 4))
```
:::

::: {.column width="2%"}
:::

::: {.column width="49%"}
## Sensitivity subset (n₁ + n₂ ≥ 40)

```{r}
#| label: ex2-tax-model-sens
#| message: false
#| warning: false

tax_model_sens_ex2 <- rma.mv(
  yi     = yi_lnM_safe,
  V      = vi_lnM_safe,
  mods   = ~ taxon.for.plot - 1 + nSE,
  random = list(~ 1 | study, ~ 1 | case.nr),
  data   = dat_sens,
  method = "REML",
  test   = "t"
)

summary(tax_model_sens_ex2)
r2_ml(tax_model_sens_ex2)

res_sens <- mod_results(
  tax_model_sens_ex2,
  group = "study",
  mod = "taxon.for.plot",
  at = list(nSE = 0)
)

orchard_plot(
  res_sens,
  mod = "taxon.for.plot",
  xlab = "lnM (SAFE)",
  group = "study",
  trunk.size = 0.6,
  branch.size = 5
) +
  scale_colour_manual(values = rep("grey20", 8)) +
  scale_fill_manual(values = rep("#117733", 4))
```
:::
::::::

### Figure 4

Comparison of overall meta-analytic estimates, taxon-level meta-regressions, and Egger-style small-study effect diagnostics for folded-normal SMD magnitude and lnM (SAFE) in the Kunc & Schmidt dataset.

```{r}
#| label: ex2-figure-4
# Combine panels into a 2 × 3 layout:
# Row 1: overall meta-analyses (absolute SMD vs lnM)
# Row 2: taxon meta-regressions (absolute SMD vs lnM)
# Row 3: Egger-style diagnostics (absolute SMD vs lnM)
(p7 + p8) / (p9 + p10) / (p11 + p12) + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 16, face = "bold"))
```

# Part II — Simulation reproduction (figures + tables)

Part II reproduces all simulation-based figures and tables reported in the manuscript using:

1)  a **pre-computed simulation summary** (one row per simulation setting), and\
2)  where needed, **replicate-level outputs** (results from each Monte Carlo replicate within a simulation setting).

A **simulation setting** (or *grid cell*) is one combination of: design × sample sizes × true lnM value (theta, θ). Within each setting, we ran several Monte Carlo replicates, computed lnM using the Delta plug-in and SAFE-BC, and summarised estimator performance.

::: callout-note
## Key quantities

| Quantity | Definition |
|:-----------------------------------|:-----------------------------------|
| $\theta$ (theta) | The *true* lnM value used to generate data in the simulation setting ($x$-axis in all plots). |
| **Bias** | $\text{mean}(\text{estimate}) - \text{true lnM}$ (computed within each setting). |
| **RMSE** | $\sqrt{\text{mean}((\text{estimate} - \text{true lnM})^2)}$ (combines bias + variability). |
| **Coverage** | The proportion of nominal 95% intervals that contain the true lnM (target is 0.95). |
| **MCSE** | Monte Carlo standard error; uncertainty in simulation summaries due to a finite number of Monte Carlo replicates. |
:::

## Figures

### Load summary results

The summary file contains one row per simulation setting (design × sample sizes × θ), including bias, RMSE, coverage, failure rates, and MCSE diagnostics computed from the Monte Carlo replicates.

```{r}
#| label: sim-load
#| code-fold: true
# Load the pre-computed simulation summary used to reproduce figures and tables results

results <- readRDS( here::here("results", "lnM_summary_SAFEfun_2025-12-21.rds"))

# Create a compact facet label for plotting 
results <- results %>%
    mutate(facet_label = paste0(design, " n1=", n1, " n2=", n2))

facet_info <- results %>% distinct(facet_label, design, n1, n2)

balanced_ind <- facet_info %>%
    filter(design == "indep", n1 == n2) %>% arrange(n1) %>% pull(facet_label)
unbalanced_ind <- facet_info %>%
    filter(design == "indep", n1 != n2) %>% arrange(n1) %>% pull(facet_label)
paired_balanced <- facet_info %>%
    filter(design == "paired") %>% arrange(n1) %>% pull(facet_label)

new_levels <- c(balanced_ind, unbalanced_ind, paired_balanced)
results <- results %>% mutate(facet_label = factor(facet_label, levels = new_levels))

strip_labeller <- function(x) gsub(" n1=", "   n1=", gsub(" n2=", "   n2=", x))

# Shared plotting theme for all simulation figures

base_theme <- theme_bw(11) +
    theme(
      legend.title = element_text(),
      legend.position = "right",
      strip.background = element_rect(fill = "grey85", colour = "grey40"),
      strip.text = element_text(face = "plain"),
      panel.grid.minor = element_blank()
)
```

### Figure 1 — Bias of point estimates

Figure 1 plots point-estimate bias for each estimator within each simulation setting. Bias is computed as
$bias = mean(estimate) - true lnM$,
so values above 0 indicate systematic overestimation and values below 0 indicate underestimation. The dashed horizontal line marks 0 bias (ideal). Each panel corresponds to one design × sample-size combination; the x-axis (theta) is the true lnM used to generate the data.

```{r}
#| label: fig1-bias
#| fig-cap: "Figure 1. Point-estimate bias (estimate − true lnM). Dashed line indicates 0 bias."
#| code-fold: true
# Construct long-format data for plotting bias curves
bias_df <- dplyr::bind_rows(
    results %>% dplyr::select(theta, facet_label, delta_bias) %>%
      dplyr::rename(bias = delta_bias) %>% dplyr::mutate(estimator = "PI"),
    results %>% dplyr::select(theta, facet_label, safe_bias) %>%
      dplyr::rename(bias = safe_bias) %>% dplyr::mutate(estimator = "SAFE")
  )

  ggplot(bias_df, aes(theta, bias, colour = estimator, group = estimator)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_line(linewidth = 0.6) +
    facet_wrap(~ facet_label, ncol = 4, nrow = 3,
               labeller = labeller(facet_label = strip_labeller)) +
    scale_color_discrete("Estimator", 
                         labels = c("PI" = "Plugin",
                                   "SAFE" = "SAFE bootstrapping"))+
    labs(x = expression(theta),
         y = "Bias (estimate − true lnM)",
         colour = "estimator") +
    base_theme + 
  theme(legend.position = "bottom")
```



### Figure 2 — Relative bias of variance estimates

Relative bias of the reported sampling variances is expressed as a percentage. In the simulation code, this is computed relative to an empirical Monte Carlo baseline variance for the SAFE-BC point estimator (computed within each setting on a common “ok” subset). The dashed line at 0% indicates an unbiased variance estimate under this baseline.

```{r}
#| label: fig2-relbias-var
#| fig-cap: "Figure 2. Relative bias of variance estimates (%). Dashed line indicates 0% (unbiased)."
#| code-fold: true

# Construct long-format data for plotting relative bias of variance estimates

rb_df <- dplyr::bind_rows(
    results %>% dplyr::select(theta, facet_label, relbias_delta) %>%
      dplyr::rename(relbias = relbias_delta) %>% dplyr::mutate(estimator = "Delta"),
    results %>% dplyr::select(theta, facet_label, relbias_safe) %>%
      dplyr::rename(relbias = relbias_safe)  %>% dplyr::mutate(estimator = "SAFE")
  )

  ggplot(rb_df, aes(theta, relbias, colour = estimator, group = estimator)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_line(linewidth = 0.6) +
    facet_wrap(~ facet_label, ncol = 4, nrow = 3,
               labeller = labeller(facet_label = strip_labeller)) +
        scale_color_discrete("Estimator", 
                         labels = c("Delta" = "Delta method derived plugin",
                                   "SAFE" = "SAFE bootstrapping"))+
    labs(x = expression(theta),
         y = "Relative bias of variance (%)",
         colour = "estimator") +
    base_theme + 
    theme(legend.position = "bottom")

```

### Figure S1 — Coverage of nominal 95% intervals

Coverage is the fraction of Monte Carlo replicates in which the nominal 95% interval contains the true lnM value. The dashed line at 0.95 is the nominal target.

```{r}
#| label: figS1-coverage
#| fig-cap: "Figure S1. Empirical coverage of nominal 95% intervals. Dashed line indicates 0.95."
#| code-fold: true
# Construct long-format data for plotting empirical coverage
cov_df <- dplyr::bind_rows(
    results %>% transmute(theta, facet_label, estimator = "PI",   cover = cover_delta),
    results %>% transmute(theta, facet_label, estimator = "SAFE", cover = cover_safe)
  )

  ggplot(cov_df, aes(theta, cover, colour = estimator, group = estimator)) +
    geom_hline(yintercept = 0.95, linetype = "dashed", colour = "grey50") +
    geom_line(linewidth = 0.6) +
    facet_wrap(~ facet_label, ncol = 4, nrow = 3,
               labeller = labeller(facet_label = strip_labeller)) +
    scale_color_discrete("Estimator", 
                         labels = c("PI" = "Plugin",
                                   "SAFE" = "SAFE bootstrapping"))+
    labs(x = expression(theta),
         y = "Empirical coverage",
         colour = "estimator") +
    base_theme + 
  theme(legend.position = "bottom")
```

### Figure S2 — RMSE of point estimates

RMSE (root mean square error) summarizes overall point-estimate accuracy by combining bias and variability. Lower RMSE indicates better performance.

```{r}
#| label: figS2-rmse
#| fig-cap: "Figure S2. RMSE of point estimates. Lower values indicate better overall accuracy."
#| code-fold: true
# Construct long-format data for plotting RMSE curves
rmse_df <- dplyr::bind_rows(
    results %>% transmute(theta, facet_label, estimator = "PI",   rmse = rmse_delta),
    results %>% transmute(theta, facet_label, estimator = "SAFE", rmse = rmse_safe)
  )

  ggplot(rmse_df, aes(theta, rmse, colour = estimator, group = estimator)) +
    geom_line(linewidth = 0.6) +
    facet_wrap(~ facet_label, ncol = 4, nrow = 3,
               labeller = labeller(facet_label = strip_labeller)) +
     scale_color_discrete("Estimator", 
                         labels = c("PI" = "Plugin",
                                   "SAFE" = "SAFE bootstrapping"))+
    labs(x = expression(theta),
         y = expression(RMSE~"("~hat(lnM)~"\u2212"~lnM~")"),
         colour = "estimator") +
    base_theme + 
  theme(legend.position = "bottom")
```

### Reading in data for Monte Carlo (MC) standard error (SE)

The chunks below read **replicate-level outputs** saved during the simulation run (`raw_runs/row_###.rds`). Each file corresponds to one simulation setting (one “grid cell”), and contains one row per Monte Carlo replicate. We use these files to reconstruct additional Monte Carlo error diagnostics (MCSE) that are easier to compute from replicate-level results. ::: callout-note Each `raw_runs/row_###.rds` file corresponds to one row of the simulation parameter grid used in the driver script (i.e., one design × sample size × θ combination). :::

```{r}
#| label: sim-mcse-load
#| code-fold: true

# Read replicate-level simulation outputs (one file per grid cell)
raw_dir <- here("raw_runs")
raw_files <- list.files(raw_dir, pattern = "^row_[0-9]{3}\\.rds$", full.names = TRUE)
if (!length(raw_files)) stop("No raw files found in: ", raw_dir)


# Extract grid-cell index from filename (e.g., row_012.rds -> 12)

row_index_from_path <- function(p) as.integer(sub("^row_([0-9]{3})\\.rds$", "\\1", basename(p)))

# Compute diagnostics for a single grid cell (one replicate-level data frame)

one_raw_diag <- function(raw_df) {
  
   # Defensive: remove replicate ID if present (not used in computations)
  if ("rep" %in% names(raw_df)) raw_df <- raw_df %>% dplyr::select(-rep)
  
  # Match the “ok_d” logic used in the simulation summary:
  # keep replicates where the Delta-method point & variance are defined and variance > 0
  ok_d <- is.finite(raw_df$delta_pt) & is.finite(raw_df$delta_var) & (raw_df$delta_var > 0)
  
  # If nothing survives the ok_d filter, return NAs
  if (!any(ok_d)) {
    return(data.frame(
      n_ok_d = 0L,
      n_ok_safe = 0L,
      RSE_VarMC_SAFEpt = NA_real_,
      sd_PI = NA_real_,
      sd_SAFE = NA_real_
    ))
  }
  
  df_ok <- raw_df[ok_d, , drop = FALSE]
  
  # Within the ok_d subset, check how many finite SAFE-BC point estimates exist
  ok_safe_pt <- is.finite(df_ok$safe_pt)
  n_ok_safe <- sum(ok_safe_pt)
  
  # Relative standard error (RSE) of the Monte Carlo variance estimate:
  # RSE(Var_hat) ≈ sqrt(2/(n-1)) for sample variance under normality
  RSE_varmc <- if (n_ok_safe >= 3) sqrt(2 / (n_ok_safe - 1)) else NA_real_
  
    # SD of point estimates (used to compute MCSE of bias later)
  # Since true_lnM is constant within a grid cell, sd(est - true) = sd(est).
  sd_PI   <- sd(df_ok$delta_pt, na.rm = TRUE)
  sd_SAFE <- sd(df_ok$safe_pt,  na.rm = TRUE)
  
  data.frame(
    n_ok_d = nrow(df_ok),
    n_ok_safe = n_ok_safe,
    RSE_VarMC_SAFEpt = RSE_varmc,
    sd_PI = sd_PI,
    sd_SAFE = sd_SAFE
  )
}
# Compute diagnostics for each grid cell (row_###.rds)
diag_list <- pbapply::pblapply(raw_files, function(f) {
  rid <- row_index_from_path(f)
  raw_df <- readRDS(f)
  d <- one_raw_diag(raw_df)
  d$row_id <- rid
  d
})

diag_df <- bind_rows(diag_list)
# Join diagnostics back onto the simulation summary
results2 <- results %>%
  mutate(row_id = seq_len(nrow(results))) %>%
  left_join(diag_df, by = "row_id")

```

### Figure S3 — MCSE of bias (from replicate-level files)

::: callout-note
**How to read MCSE plots (Figures S3–S4).**\
MCSE is simulation uncertainty, not estimator uncertainty. It measures how precisely the simulation has estimated quantities like bias and mean(variance). Lower MCSE implies more stable simulation summaries.
:::

Figure S3 shows the **Monte Carlo standard error (MCSE)** of the estimated bias within each simulation setting. Here, MCSE quantifies how much the **estimated bias** would vary if we re-ran the Monte Carlo simulation again with the same number of replicates.

We compute this from replicate-level outputs using the approximation:

`MCSE(bias) ≈ sd(estimate) / sqrt(n_eff)`

where `n_eff` is the number of usable Monte Carlo replicates in that setting (after applying the same validity filters used in the simulation summary). Smaller MCSE values indicate more precise bias estimates.

```{r}
#| label: figS3-mcse-bias
#| fig-cap: "Figure S3. MCSE of bias (per grid cell) reconstructed from replicate-level files in raw_runs."
#| code-fold: true
# MCSE(bias) approximation:
#   MCSE(bias) ≈ sd(estimator) / sqrt(n_eff),
# where n_eff is the number of usable replicates in each grid cell.

results2 <- results2 %>%
  mutate(
    mcse_bias_PI_disk   = ifelse(is.finite(sd_PI)   & n_ok_d >= 2,    sd_PI   / sqrt(n_ok_d),     NA_real_),
    mcse_bias_SAFE_disk = ifelse(is.finite(sd_SAFE) & n_ok_safe >= 2, sd_SAFE / sqrt(n_ok_safe),  NA_real_)
  )

mc_long <- results2 %>%
  transmute(
    theta, facet_label,
    mcse_bias_PI   = mcse_bias_PI_disk,
    mcse_bias_SAFE = mcse_bias_SAFE_disk
  ) %>%
  pivot_longer(
    cols = c(mcse_bias_PI, mcse_bias_SAFE),
    names_to = "metric", values_to = "value"
  )

p_mc <- ggplot(mc_long, aes(theta, value, colour = metric, group = metric)) +
  geom_line() +
  facet_wrap(~ facet_label, ncol = 4) +
  scale_color_discrete("Estimator", 
                         labels = c("mcse_bias_PI" = "Plugin",
                                   "mcse_bias_SAFE" = "SAFE bootstrapping"))+
  labs(
    x = expression(theta),
    y = "MCSE of bias",
    title = "Monte-Carlo error diagnostics: MCSE of bias (per grid cell; from raw runs)"
  ) +
  base_theme + 
  theme(legend.position = "bottom")

print(p_mc)

```

### Figure S4 — MCSE of mean(variance estimate) (from summary)

Figure S4 shows the MCSE of the **mean reported sampling variance** for each estimator within each simulation setting. This is a precision diagnostic for the simulation itself: it indicates how much uncertainty remains in the simulation summary of `mean_var_delta` and `mean_var_safe` due to finite Monte Carlo replicates.

In practice, smaller MCSE values mean the simulation is sufficiently large to estimate average variance behavior reliably for that setting; larger values indicate that more Monte Carlo replicates would be needed to stabilize the mean variance estimate.

```{r}
#| label: figS4-mcse-varmean
#| fig-cap: "Figure S4. MCSE of mean variance estimates (from the simulation summary)."
#| code-fold: true

# Construct long-format data for plotting MCSE of mean variance estimates
var_mcse_df <- results %>%
  transmute(theta, facet_label,
            PI_var_MCSE   = mcse_varbar_delta,
            SAFE_var_MCSE = mcse_varbar_safe) %>%
  pivot_longer(cols = c(PI_var_MCSE, SAFE_var_MCSE),
               names_to = "estimator", values_to = "mcse") %>%
  mutate(estimator = recode(estimator,
                            PI_var_MCSE   = "PI (delta-var)",
                            SAFE_var_MCSE = "SAFE (safe-var)"))

p_var_mcse <- ggplot(var_mcse_df, aes(theta, mcse, colour = estimator, group = estimator)) +
  geom_line() +
  facet_wrap(~ facet_label, ncol = 4) +
  labs(x = expression(theta),
       y = "MCSE of mean(variance estimate)",
       colour = NULL,
       title = "Monte-Carlo error of variance estimators (MCSE of the mean)") +
  base_theme + 
  theme(legend.position = "bottom")

print(p_var_mcse)
```

## Tables for the supplement

### Table S1 — Method feasibility and stability

Table S1 summarizes feasibility and stability diagnostics for lnM estimation across simulation settings. Values are summarized across theta within each design × (n1, n2) facet. Medians represent typical behavior across theta, while maxima highlight worst-case behavior near boundary regions.

| Variable Name | Definition |
|:-----------------------------------|:-----------------------------------|
| `θ_min`, `θ_max` | Minimum and maximum true lnM values evaluated in the simulation grid for a given design × sample-size combination. |
| `delta_fail_median` | Median proportion of simulation replicates (across $\theta$) for which the Delta plug-in lnM estimator was undefined (e.g., due to boundary violations). |
| `safe_fail_median` | Median proportion of simulation replicates for which SAFE failed to return a usable lnM estimate. |
| `SAFE_ok_median` | Median proportion of replicates in which SAFE returned a valid (accepted) bootstrap estimate. |
| `delta_cap_median` | Median proportion of replicates in which the Delta-method sampling variance was capped for numerical stability. |
| `delta_fail_max`, `safe_fail_max`, `delta_cap_max` | Worst-case (maximum) values of the corresponding metrics across $\theta$, highlighting boundary regions where estimator behavior deteriorates. |

::: {.callout-note appearance="simple"}
```{r}
#| label: tabS1-feasibility
#| code-fold: true

tab1_stability <- results %>%
  group_by(design, n1, n2, facet_label) %>%
  summarise(
    # Range of θ (theta) values evaluated in each facet
    theta_min = min(theta, na.rm = TRUE),
    theta_max = max(theta, na.rm = TRUE),
    
    # Median failure rates across θ
    delta_fail_median = median(delta_fail_prop, na.rm = TRUE),
    safe_fail_median  = median(safe_fail_prop,  na.rm = TRUE),
    
    # Median SAFE convergence / acceptance rate
    SAFE_ok_median    = median(SAFE_ok_rate, na.rm = TRUE),
    
    # Median rate of Delta variance capping
    delta_cap_median  = median(delta_cap_rate, na.rm = TRUE),
    
    # Worst-case behaviour across θ
    delta_fail_max = max(delta_fail_prop, na.rm = TRUE),
    safe_fail_max  = max(safe_fail_prop,  na.rm = TRUE),
    delta_cap_max  = max(delta_cap_rate,  na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  arrange(factor(facet_label, levels = levels(results$facet_label))) %>%
  mutate(
    design = as.character(design),
    facet_label = as.character(facet_label)
  )

kable(tab1_stability, digits = 3,
      caption = "Table S1. Feasibility / stability diagnostics summarised across θ within each design × (n1,n2) facet.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))%>%
  kableExtra::scroll_box(width = "100%", height = "400px")
```
:::

### Table S2 — SAFE bootstrap acceptance (kept/tried)

Table S2 summarizes the SAFE bootstrap acceptance rate (“kept / tried”) across theta within each facet. Quantiles show how acceptance varies across theta, including worst-case regions where accepted draws are rare.

| Variable Name | Definition |
|:-----------------------------------|:-----------------------------------|
| `q0`, `q25`, `q50`, `q75`, `q100` | Minimum, 25th percentile, median, 75th percentile, and maximum SAFE bootstrap acceptance rates across values of $\theta$ within each facet. |
| `mean` | Mean SAFE bootstrap acceptance rate across $\theta$. |

::: {.callout-note appearance="simple"}
```{r}
#| label: tabS2-safe-acceptance
#| code-fold: true

tab2_accept <- results %>%
  group_by(design, n1, n2, facet_label) %>%
  summarise(
    q0  = quantile(boot_accept_prop, 0.00, na.rm = TRUE),
    q25 = quantile(boot_accept_prop, 0.25, na.rm = TRUE),
    q50 = quantile(boot_accept_prop, 0.50, na.rm = TRUE),
    q75 = quantile(boot_accept_prop, 0.75, na.rm = TRUE),
    q100= quantile(boot_accept_prop, 1.00, na.rm = TRUE),
    mean = mean(boot_accept_prop, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(factor(facet_label, levels = levels(results$facet_label)))

kable(tab2_accept, digits = 4,
      caption = "Table S2. SAFE acceptance rate (kept/tried) summarised across θ for each facet. Quantiles highlight worst-case acceptance.") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed"))%>%
  kableExtra::scroll_box(width = "100%", height = "400px")
```

Acceptance rate is defined as the proportion of bootstrap resamples retained after SAFE’s feasibility and stability checks. Lower values indicate reduced effective bootstrap sample sizes, typically occurring near parameter boundaries or under small sample sizes.
:::

### Table S3 — MCSE diagnostics

Table S3 summarizes Monte Carlo precision. Values are the maximum MCSE observed across theta within each facet, giving a conservative measure of simulation uncertainty.

| Variable Name | Definition |
|:-----------------------------------|:-----------------------------------|
| `mcse_bias_PI_max` | Maximum Monte Carlo standard error of the Delta-method lnM bias estimate across $\theta$. |
| `mcse_bias_SAFE_max` | Maximum Monte Carlo standard error of the SAFE lnM bias estimate across $\theta$. |
| `mcse_var_PI_max` | Maximum Monte Carlo standard error of the mean Delta-method variance estimate across $\theta$. |
| `mcse_var_SAFE_max` | Maximum Monte Carlo standard error of the mean SAFE variance estimate across $\theta$. |

: Column definitions for Monte Carlo standard errors {#tbl-mcse-defs}

::: {.callout-note appearance="simple"}
```{r}
#| label: tabS3-mcse
#| code-fold: true

has_mcse_bias <- all(c("mcse_bias_delta","mcse_bias_safe") %in% names(results))

tab3_mcse <- results %>%
  group_by(design, n1, n2, facet_label) %>%
  summarise(
    # bias MCSE (optional)
    mcse_bias_PI_max   = if (has_mcse_bias) max(mcse_bias_delta, na.rm = TRUE) else NA_real_,
    mcse_bias_SAFE_max = if (has_mcse_bias) max(mcse_bias_safe,  na.rm = TRUE) else NA_real_,
    
    # variance MCSE (these are in your results)
    mcse_var_PI_max    = max(mcse_varbar_delta, na.rm = TRUE),
    mcse_var_SAFE_max  = max(mcse_varbar_safe,  na.rm = TRUE),
    
    .groups = "drop"
  ) %>%
  arrange(factor(facet_label, levels = levels(results$facet_label)))

kable(tab3_mcse, digits = 5,
      caption = "Table S3. Monte-Carlo precision (max MCSE over θ within each facet).") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped","hover","condensed"))%>%
  kableExtra::scroll_box(width = "100%", height = "400px")

```

MCSE values quantify simulation uncertainty due to a finite number of Monte Carlo replicates; smaller values indicate more precise simulation estimates.
:::

## References

::: {#refs}
:::

# Session info

```{r}
#| label: session-info
sessionInfo()
```
